---
title: "GSS EDA - health and covariates"
author: "Christine Lucille Kuryla"
date: "2024-09-27"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gridExtra)
```

Take a look at GSS dataset, what's available, some trends, etc. 

The first variable of interest is "health", which will be the main subject of our analysis. 

https://gssdataexplorer.norc.org/variables/437/vshow

Question on survey: "Would you say your own health, in general, is excellent, good, fair, or  poor?"

# Fetch GSS data 

```{r fetch_data, eval = FALSE}

# Feel free to modify to play with more covariates and variables.

#install.packages('gssr', repos =  c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))
# install.packages('gssrdoc', repos = c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))

library(gssr)
library(gssrdoc)

data("gss_all"). # this file is big!

data_gss <- as.data.frame(gss_all) %>% 
  select("year",      # year of survey
         "cohort",    # birthyear
         "age",       # age at time of survey
         "health",    # self-rated health
         "sex",       # sex
         "happy",     # self-rated happiness
         "life",      # is life exciting or dull
         "educ",      # years of education
         "polviews",  # 1 extremely liberal, 4 moderate, 7 extremely conservative
         "class",    # 1 lower, 2 middle, 3 working, 4 upper, 5 no class
         "satfin"    # 1 pretty well satisfied, 2 more or less satisfied, 3 not satisfied at all
         )

write_csv(data_gss, "data/extracted_gss_variables.csv")

```

```{r load_and_clean_data}
data_gss <- read_csv("data/extracted_gss_variables.csv") %>% 
  filter(cohort != 9999) %>% 
  na.omit() %>% 
  mutate(health = 5 - health)  %>%  # reverse the coding so it's more intuitive (higher number for excellent, lower number for poor)
  mutate(happy = 4 - happy) %>% # same
  mutate(life = 4 - life) %>% # reverse again, these variables tend to be unintuitively ordered!!!
  mutate(satfin = 4 - satfin) 
```

## Peek at df
```{r head}

head(data_gss)

```


## Life expectancy data

```{r, eval = FALSE}

# https://population.un.org/wpp/Download/Standard/CSV/
data_le <- read_csv("big_data/WPP2024_Demographic_Indicators_Medium.csv")
data_le <- data_le %>% 
  filter(Location == "United States of America")

# Just get the years of interest for merging with the gss data
year_range <- seq(min(data_gss$year), max(data_gss$year))
data_le <- data_le %>% 
  filter(Time %in% year_range)

write_csv(data_le, "data/un_life_expectancy_etc.csv")

```


# Histograms

```{r eda_general}

# Tidyverse and flexible number of histograms

library(patchwork)

# Create list to store ggplot objects
plot_list <- list()

# Loop through column names and create plots
for (var in colnames(data_gss)) {
  if (is.numeric(data_gss[[var]])) {  # Only create histograms for numeric columns
    p <- ggplot(data_gss, aes(x = .data[[var]])) +  # Use .data[[var]] for tidy evaluation
      geom_histogram(bins = 20, fill = "pink", color = "hotpink") +
      theme_minimal()
    
    plot_list[[var]] <- p
  }
}

# Combine all plots using patchwork and add a title
combined_plot <- wrap_plots(plot_list, ncol = 3) +
  plot_annotation(title = "Histograms of Survey Responses")

# Display the combined plot
print(combined_plot)



```


# Correlation Heatmap

```{r}
correlation_auto <- cor(data_gss)
knitr::kable(correlation_auto)

# Create a heatmap

#upper_tri <- matrixcalc::upper.triangle(correlation_auto)
#melted_cormat <- reshape2::melt(upper_tri, na.rm = TRUE)

melted_cormat <- reshape2::melt(cor(data_gss), na.rm = TRUE)

ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
     hjust = 1))+
 coord_fixed() +
  geom_text(aes(Var2, Var1, label = if_else(value != 0, as.character(round(value, digits = 2)), " ")))
```

# Mean health over A/P/C

```{r over_apc}

# Mean health over time

data_gss %>% 
  group_by(year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health)) +
  geom_line() +
  labs(title = "Self-Rated Health Each Year")

data_gss %>% 
  group_by(age) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line()

data_gss %>% 
  group_by(cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = cohort, y = mean_health)) +
  geom_line()

```

# Some more EDA

```{r}

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% 
  group_by(year, cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = cohort)) +
  geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% 
  group_by(age, cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health, color = cohort)) +
  geom_line()



```

# Happiness and health

```{r happiness_and_health}

data_gss %>% 
  ggplot(aes(x = health, y = happy)) +
  geom_smooth(method = "lm")

data_gss %>% 
  mutate(year = cut(year, breaks = 6)) %>% 
  mutate(age = cut(age, breaks = 6)) %>% 
  ggplot(aes(x = health, y = happy, color = age)) +
  facet_wrap(~ year) +
  geom_smooth(method = "lm")

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% 
  mutate(year = cut(year, breaks = 6)) %>% 
  ggplot(aes(x = health, y = happy, color = cohort)) +
  facet_wrap(~ year) +
  geom_smooth(method = "lm")

######data_gss %>% 
  # group_by(year) %>% 
  # summarize(mean_health = mean(health)) %>% 
  # ggplot(aes(x = cohort, y = mean_health)) +
  # geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% 
#  mutate(year = cut(year, breaks = 6)) %>% 
#  mutate(age = cut(year, breaks = 6)) %>% 
  ggplot(aes(x = age, y = health, color = cohort)) +
#  facet_wrap(~ year) +
  geom_smooth(method = "lm")

```

