---
title: "Panel Data NLSY97"
author: "Christine Lucille Kuryla"
date: "2024-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


This analysis uses panel data from NLSY97 (National Longitudinal Survey Youth 1997). More information and data can be found here: https://www.bls.gov/nls/nlsy97.htm

A total of 6 waves have the variables of interest, so we do not explore first differences, just fixed and random effects.


Outcome variable: `health`
Our outcome variable of interest is self-rated health. In this dataset, the variable has been recoded and named health and ranges from 1 (Poor) to 5 (Excellent) as detailed below. 

Main predictor: `watched_tv`
Our main predictor variable of interest is how often the participant watches TV. The interpretation ranges from 1 (No TV) to 6 (10+ hours per week). The values are described below.  \n
1 - "None" \n
2 - "Less than 1 hour a week" \n
3 - "1 to 3 hours a week" \n
4 - "4 to 6 hours a week" \n
5 - "7 to 9 hours a week" \n
6 - "10 hours or more a week" \n

Other predictor: `felt_happy` \n
How often the respondent felt happy in the last month. \n
This variable has been recoded such that it ranges from: \n
4 - "All of the time" \n
3 - "Most of the time" \n
2 - "Some of the time" \n
1 - "None of the time" \n

Other predictor: `depressed`
How often the respondent was depressed in the last month. \n
This variable has been recoded such that it ranges from: \n
4 - "All of the time" \n
3 - "Most of the time" \n
2 - "Some of the time" \n
1 - "None of the time" \n


# Variables of interest description

How would you rate your health?

(Note these have been recoded)

5 - "Excellent"

4 - "Very good"

3 - "Good"

2 - "Fair"

1 - "Poor"


How often felt happy in the last month:

How often felt down or blue in the last month:

How often felt calm in the last month:

How often felt like a nervous person in the last month:

How often felt depressed in the last month:

(Note these have been recoded from original data)

4 - "All of the time"

3 - "Most of the time"

2 - "Some of the time"

1 - "None of the time"


Watches TV

1 - "None"

2 - "Less than 1 hour a week"

3 - "1 to 3 hours a week"

4 - "4 to 6 hours a week"

5 - "7 to 9 hours a week"

6 - "10 hours or more a week"


Uses computer

1 - "Less than 2 hours per week"

2 - "3 to 10 hours a week"

3 - "11 to 20 hours a week"

4 - "21 to 30 hours a week"

5 - "31 to 40 hours a week"

6 - "More than 40 hours a week"

# Wrangle
```{r import_and_format_data, eval = FALSE, include = FALSE}

new_data <- read.table('big_data/panel/more_nls97_0/more_nls97_0.dat', sep=' ')
names(new_data) <- c('R0000100',
'R0320600',
'R0320800',
'R0320900',
'R0321000',
'R0321100',
'R0321200',
'R0536300',
'R0536401',
'R0536402',
'R0536600',
'R1193900',
'R1194000',
'R1194100',
'R1200100',
'R1200200',
'R1200300',
'R1201300',
'R1210400',
'R1217500',
'R1235800',
'R1302400',
'R1302500',
'R1302600',
'R1302700',
'R1482600',
'R2164000',
'R2363400',
'R2363500',
'R2553400',
'R2553500',
'R2558800',
'R2569500',
'R2576800',
'R3481900',
'R3670400',
'R3670500',
'R3876200',
'R3876300',
'R3880300',
'R3891500',
'R3899100',
'R4880100',
'R4893600',
'R4893700',
'R4893800',
'R4893900',
'R4894000',
'R5120900',
'R5121000',
'R5453600',
'R5453700',
'R5459400',
'R5473500',
'R5484100',
'R6497500',
'R6850300',
'R6850400',
'R7215900',
'R7216000',
'R7222400',
'R7237200',
'R7248400',
'R7295600',
'S0920800',
'S0920900',
'S0921000',
'S0921100',
'S0921200',
'S1081000',
'S1081100',
'S1225000',
'S1225100',
'S1225200',
'S1225300',
'S1225400',
'S1225500',
'S1225600',
'S1225700',
'S1225800',
'S1531300',
'S1531400',
'S1535500',
'S1552300',
'S1552500',
'S1564300',
'S2000900',
'S2001000',
'S2005400',
'S2022100',
'S2022300',
'S2034400',
'S3154600',
'S3154700',
'S3302500',
'S3801000',
'S3801100',
'S3805700',
'S3823000',
'S3823200',
'S3835800',
'S4681900',
'S4682000',
'S4682100',
'S4682200',
'S4682300',
'S4812800',
'S4812900',
'S4919500',
'S5400900',
'S5401000',
'S5405600',
'S5423200',
'S5436300',
'S6511400',
'S6511500',
'S6661100',
'S7501100',
'S7501200',
'S7506100',
'S7509600',
'S7525300',
'S7537100',
'S8332300',
'S8332400',
'S8332500',
'S8332600',
'S8332700',
'S8509800',
'S8509900',
'S8644200',
'S8646800',
'S8647900',
'T0008400',
'T0008500',
'T0009400',
'T0012900',
'T0025600',
'T0033700',
'T0060700',
'T0902900',
'T0903000',
'T1049500',
'T1049600',
'T1049700',
'T1049800',
'T1049900',
'T1050000',
'T1050100',
'T1050200',
'T1050300',
'T1068300',
'T2011000',
'T2011100',
'T2012100',
'T2015800',
'T2020500',
'T2021300',
'T2043400',
'T2782600',
'T2782700',
'T2782800',
'T2782900',
'T2783000',
'T3014700',
'T3014800',
'T3144600',
'T3144800',
'T3144900',
'T3145000',
'T3145100',
'T3145200',
'T3145300',
'T3145400',
'T3145500',
'T3162400',
'T3601400',
'T3601500',
'T3602100',
'T3606100',
'T3611200',
'T3612000',
'T3645400',
'T4562200',
'T4562300',
'T4562400',
'T4565400',
'T4565500',
'T4565600',
'T4565700',
'T4565800',
'T5201300',
'T5201400',
'T5202300',
'T5206500',
'T5211600',
'T5211900',
'T6143700',
'T6143800',
'T6143900',
'T6144000',
'T6144100',
'T6206400',
'T6206500',
'T6206600',
'T6209600',
'T6209700',
'T6209800',
'T6209900',
'T6210000',
'T6651200',
'T6651300',
'T6652100',
'T6656300',
'T6663000',
'T6663300',
'T7703800',
'T7703900',
'T7704000',
'T7707000',
'T7707100',
'T7707200',
'T7707300',
'T7707400',
'T8123500',
'T8123600',
'T8123700',
'T8128700',
'T8134200',
'T8134500',
'T9093100',
'U0001700',
'U0001800',
'U0001900',
'U0008600',
'U0014700',
'U0015000',
'U1030700',
'U1030800',
'U1030900',
'U1031000',
'U1031100',
'U1096500',
'U1096600',
'U1096700',
'U1099700',
'U1099800',
'U1099900',
'U1100000',
'U1100100',
'U1838400',
'U1838500',
'U1838600',
'U1852500',
'U1853200',
'U2962600',
'U2962800',
'U2962900',
'U2963000',
'U2963200',
'U2963300',
'U2964800',
'U2964900',
'U2965000',
'U2965100',
'U2965200',
'U3437800',
'U3437900',
'U3438000',
'U3451600',
'U3453600',
'U4368000',
'U4369900',
'U4370000',
'U4370100',
'U4370200',
'U4370300',
'U4942900',
'U4943000',
'U4943100',
'U4954700',
'U4956900',
'U5860700',
'U5867900',
'U5868000',
'U5868001',
'U5868002',
'U5868003',
'U5868004',
'U5868100',
'Z0397600',
'Z0397700',
'Z0530200')


# Handle missing values

new_data[new_data == -1] = NA  # Refused
new_data[new_data == -2] = NA  # Dont know
new_data[new_data == -3] = NA  # Invalid missing
new_data[new_data == -4] = NA  # Valid missing
new_data[new_data == -5] = NA  # Non-interview


# If there are values not categorized they will be represented as NA

vallabels = function(data) {
  data$R0320600 <- factor(data$R0320600,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$R0320800 <- factor(data$R0320800,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7"))
  data$R0320900 <- factor(data$R0320900,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7"))
  data$R0321000 <- factor(data$R0321000,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7"))
  data$R0321100 <- factor(data$R0321100,
levels=c(0.0,1.0,999.0),
labels=c("NEVER",
"ENTER A PERCENT  (Enter percent on next screen.)",
"Do not ride in a car"))
  data$R0536300 <- factor(data$R0536300,
levels=c(0.0,1.0,2.0),
labels=c("No Information",
"Male",
"Female"))
  data$R0536401 <- factor(data$R0536401,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,11.0,12.0),
labels=c("1: January",
"2: February",
"3: March",
"4: April",
"5: May",
"6: June",
"7: July",
"8: August",
"9: September",
"10: October",
"11: November",
"12: December"))
  data$R1200300 <- factor(data$R1200300,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$R1201300 <- factor(data$R1201300,
levels=c(1.0,2.0,3.0),
labels=c("Citizen, born in the U.S.",
"Unknown, not born in U.S.",
"Unknown, can't determine birthplace"))
  data$R1210400 <- factor(data$R1210400,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("not in MSA",
"in MSA, not in central city",
"in MSA, in central city",
"in MSA, not known",
"not in country"))
  data$R1217500 <- factor(data$R1217500,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$R1235800 <- factor(data$R1235800,
levels=c(0.0,1.0),
labels=c("Oversample",
"Cross-sectional"))
  data$R1302400 <- factor(data$R1302400,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,11.0,12.0,13.0,14.0,15.0,16.0,17.0,18.0,19.0,20.0,95.0),
labels=c("NONE",
"1ST GRADE",
"2ND GRADE",
"3RD GRADE",
"4TH GRADE",
"5TH GRADE",
"6TH GRADE",
"7TH GRADE",
"8TH GRADE",
"9TH GRADE",
"10TH GRADE",
"11TH GRADE",
"12TH GRADE",
"1ST YEAR COLLEGE",
"2ND YEAR COLLEGE",
"3RD YEAR COLLEGE",
"4TH YEAR COLLEGE",
"5TH YEAR COLLEGE",
"6TH YEAR COLLEGE",
"7TH YEAR COLLEGE",
"8TH YEAR COLLEGE OR MORE",
"UNGRADED"))
  data$R1302500 <- factor(data$R1302500,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,11.0,12.0,13.0,14.0,15.0,16.0,17.0,18.0,19.0,20.0,95.0),
labels=c("NONE",
"1ST GRADE",
"2ND GRADE",
"3RD GRADE",
"4TH GRADE",
"5TH GRADE",
"6TH GRADE",
"7TH GRADE",
"8TH GRADE",
"9TH GRADE",
"10TH GRADE",
"11TH GRADE",
"12TH GRADE",
"1ST YEAR COLLEGE",
"2ND YEAR COLLEGE",
"3RD YEAR COLLEGE",
"4TH YEAR COLLEGE",
"5TH YEAR COLLEGE",
"6TH YEAR COLLEGE",
"7TH YEAR COLLEGE",
"8TH YEAR COLLEGE OR MORE",
"UNGRADED"))
  data$R1302600 <- factor(data$R1302600,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,11.0,12.0,13.0,14.0,15.0,16.0,17.0,18.0,19.0,20.0,95.0),
labels=c("NONE",
"1ST GRADE",
"2ND GRADE",
"3RD GRADE",
"4TH GRADE",
"5TH GRADE",
"6TH GRADE",
"7TH GRADE",
"8TH GRADE",
"9TH GRADE",
"10TH GRADE",
"11TH GRADE",
"12TH GRADE",
"1ST YEAR COLLEGE",
"2ND YEAR COLLEGE",
"3RD YEAR COLLEGE",
"4TH YEAR COLLEGE",
"5TH YEAR COLLEGE",
"6TH YEAR COLLEGE",
"7TH YEAR COLLEGE",
"8TH YEAR COLLEGE OR MORE",
"UNGRADED"))
  data$R1302700 <- factor(data$R1302700,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,11.0,12.0,13.0,14.0,15.0,16.0,17.0,18.0,19.0,20.0,95.0),
labels=c("NONE",
"1ST GRADE",
"2ND GRADE",
"3RD GRADE",
"4TH GRADE",
"5TH GRADE",
"6TH GRADE",
"7TH GRADE",
"8TH GRADE",
"9TH GRADE",
"10TH GRADE",
"11TH GRADE",
"12TH GRADE",
"1ST YEAR COLLEGE",
"2ND YEAR COLLEGE",
"3RD YEAR COLLEGE",
"4TH YEAR COLLEGE",
"5TH YEAR COLLEGE",
"6TH YEAR COLLEGE",
"7TH YEAR COLLEGE",
"8TH YEAR COLLEGE OR MORE",
"UNGRADED"))
  data$R1482600 <- factor(data$R1482600,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Black",
"Hispanic",
"Mixed Race (Non-Hispanic)",
"Non-Black / Non-Hispanic"))
  data$R2164000 <- factor(data$R2164000,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$R2363400 <- factor(data$R2363400,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$R2363500 <- factor(data$R2363500,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,97.0),
labels=c("LIVE WITH PARENT(S)",
"LIVE WITH SPOUSE'S/PARTNER'S PARENT(S)",
"HOUSING IS PART OF JOB COMPENSATION; LIVE-IN SERVANT; HOUSEKEEPER; GARDENER; FARM LABORER",
"HOUSING IS A GIFT PAID FOR BY AN HU RESIDENT OTHER THAN R OR SPOUSE/PARTNER",
"HOUSING IS A GIFT PAID FOR BY A FRIEND OR RELATIVE OUTSIDE OF THE HU",
"HOUSING PAID FOR BY  A GOVERNMENT AGENCY/WELFARE/CHARITABLE INSTITUTION",
"SOLD HOME, NOT MOVED OUT OF IT YET",
"LIVING IN HOUSE WHICH R WILL INHERIT; ESTATE IN PROGRESS",
"LIVING IN TEMPORARY QUARTERS (GARAGE, SHED) WHILE HOME IS UNDER CONSTRUCTION",
"OTHER"))
  data$R2558800 <- factor(data$R2558800,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$R2569500 <- factor(data$R2569500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("not in MSA",
"in MSA, not in central city",
"in MSA, in central city",
"in MSA, not known",
"not in country"))
  data$R2576800 <- factor(data$R2576800,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$R3481900 <- factor(data$R3481900,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$R3670400 <- factor(data$R3670400,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$R3670500 <- factor(data$R3670500,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,97.0),
labels=c("LIVE WITH PARENT(S)",
"LIVE WITH SPOUSE'S/PARTNER'S PARENT(S)",
"HOUSING IS PART OF JOB COMPENSATION; LIVE-IN SERVANT; HOUSEKEEPER; GARDENER; FARM LABORER",
"HOUSING IS A GIFT PAID FOR BY AN HU RESIDENT OTHER THAN R OR SPOUSE/PARTNER",
"HOUSING IS A GIFT PAID FOR BY A FRIEND OR RELATIVE OUTSIDE OF THE HU",
"HOUSING PAID FOR BY  A GOVERNMENT AGENCY/WELFARE/CHARITABLE INSTITUTION",
"SOLD HOME, NOT MOVED OUT OF IT YET",
"LIVING IN HOUSE WHICH R WILL INHERIT; ESTATE IN PROGRESS",
"LIVING IN TEMPORARY QUARTERS (GARAGE, SHED) WHILE HOME IS UNDER CONSTRUCTION",
"OTHER"))
  data$R3880300 <- factor(data$R3880300,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$R3891500 <- factor(data$R3891500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("not in MSA",
"in MSA, not in central city",
"in MSA, in central city",
"in MSA, not known",
"not in country"))
  data$R3899100 <- factor(data$R3899100,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$R4880100 <- factor(data$R4880100,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$R4893600 <- factor(data$R4893600,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$R4893700 <- factor(data$R4893700,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$R4893800 <- factor(data$R4893800,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$R4893900 <- factor(data$R4893900,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$R4894000 <- factor(data$R4894000,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$R5120900 <- factor(data$R5120900,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$R5121000 <- factor(data$R5121000,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,97.0),
labels=c("LIVE WITH PARENT(S)",
"LIVE WITH SPOUSE'S/PARTNER'S PARENT(S)",
"HOUSING IS PART OF JOB COMPENSATION; LIVE-IN SERVANT; HOUSEKEEPER; GARDENER; FARM LABORER",
"HOUSING IS A GIFT PAID FOR BY AN HU RESIDENT OTHER THAN R OR SPOUSE/PARTNER",
"HOUSING IS A GIFT PAID FOR BY A FRIEND OR RELATIVE OUTSIDE OF THE HU",
"HOUSING PAID FOR BY  A GOVERNMENT AGENCY/WELFARE/CHARITABLE INSTITUTION",
"SOLD HOME, NOT MOVED OUT OF IT YET",
"LIVING IN HOUSE WHICH R WILL INHERIT; ESTATE IN PROGRESS",
"LIVING IN TEMPORARY QUARTERS (GARAGE, SHED) WHILE HOME IS UNDER CONSTRUCTION",
"OTHER"))
  data$R5459400 <- factor(data$R5459400,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$R5473500 <- factor(data$R5473500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("not in MSA",
"in MSA, not in central city",
"in MSA, in central city",
"in MSA, not known",
"not in country"))
  data$R5484100 <- factor(data$R5484100,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$R6497500 <- factor(data$R6497500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$R6850300 <- factor(data$R6850300,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$R6850400 <- factor(data$R6850400,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,97.0),
labels=c("Live with Parent(s)",
"Live with spouse's/partner's parents",
"Housing is part of job compensation; live-in servant; gardener; farm laborer",
"Housing is a gift paid for by an HU resident other than R or spouse/partner",
"Housing paid for by a government agency/welfare/charitable institution",
"Housing paid for by a government agency/welfare/charitable institution",
"Sold home, not moved out of it yet",
"Living in house which R will inherit; estate in progress",
"Living in temporary quarters (garage, shed) while home is under construction",
"Other"))
  data$R7222400 <- factor(data$R7222400,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$R7237200 <- factor(data$R7237200,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("not in MSA",
"in MSA, not in central city",
"in MSA, in central city",
"in MSA, not known",
"not in country"))
  data$R7248400 <- factor(data$R7248400,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$R7295600 <- factor(data$R7295600,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$S0920800 <- factor(data$S0920800,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S0920900 <- factor(data$S0920900,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S0921000 <- factor(data$S0921000,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S0921100 <- factor(data$S0921100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S0921200 <- factor(data$S0921200,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S1081000 <- factor(data$S1081000,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$S1081100 <- factor(data$S1081100,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,97.0),
labels=c("Live with parent(s)",
"Live with spouse's/partner's parent(s)",
"Housing is part of job compensation; live-in servant; housekeeper; gardener; farm laborer",
"Housing is a gift paid for by an HU resident other than R or spouse/partner",
"Housing is a gift paid for by a friend or relative outside of the HU",
"Housing paid for by a government agency/welfare/charitable institution",
"Sold home, not moved out of it yet",
"Living in house which R will inherit; estate in progress",
"Living in temporary quarters (garage, shed) while home is under construction",
"Other"))
  data$S1225000 <- factor(data$S1225000,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$S1225100 <- factor(data$S1225100,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat fruit",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$S1225200 <- factor(data$S1225200,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat vegetables",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$S1225400 <- factor(data$S1225400,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("None",
"Less than 1 hour a week",
"1 to 3 hours a week",
"4 to 6 hours a week",
"7 to 9 hours a week",
"10 hours or more a week"))
  data$S1225500 <- factor(data$S1225500,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Less than 2 hours per week",
"3 to 10 hours a week",
"11 to 20 hours a week",
"21 to 30 hours a week",
"31 to 40 hours a week",
"More than 40 hours a week"))
  data$S1225700 <- factor(data$S1225700,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I DID NOT DRIVE IN THE PAST 30 DAYS"))
  data$S1225800 <- factor(data$S1225800,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I HAVE NOT BEEN A PASSENGER IN THE PAST 30 DAYS"))
  data$S1535500 <- factor(data$S1535500,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$S1552300 <- factor(data$S1552300,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$S1552500 <- factor(data$S1552500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("not in MSA",
"in MSA, not in central city",
"in MSA, in central city",
"in MSA, not known",
"not in country"))
  data$S1564300 <- factor(data$S1564300,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$S2005400 <- factor(data$S2005400,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$S2022100 <- factor(data$S2022100,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$S2022300 <- factor(data$S2022300,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("not in MSA",
"in MSA, not in central city",
"in MSA, in central city",
"in MSA, not known",
"not in country"))
  data$S2034400 <- factor(data$S2034400,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$S3154600 <- factor(data$S3154600,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$S3154700 <- factor(data$S3154700,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,97.0),
labels=c("Live with parent(s)",
"Live with spouse's/partner's parent(s)",
"Housing is part of job compensation; live-in servant; housekeeper; gardener; farm laborer",
"Housing is a gift paid for by an HU resident other than R or spouse/partner",
"Housing is a gift paid for by a friend or relative outside of the HU",
"Housing paid for by a government agency/welfare/charitable institution",
"Sold home, not moved out of it yet",
"Living in house which R will inherit; estate in progress",
"Living in temporary quarters (garage, shed) while home is under construction",
"Other"))
  data$S3302500 <- factor(data$S3302500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$S3805700 <- factor(data$S3805700,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$S3823000 <- factor(data$S3823000,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$S3823200 <- factor(data$S3823200,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("not in CBSA",
"in CBSA, not in central city",
"in CBSA, in central city",
"in CBSA, not known",
"not in country"))
  data$S3835800 <- factor(data$S3835800,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$S4681900 <- factor(data$S4681900,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S4682000 <- factor(data$S4682000,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S4682100 <- factor(data$S4682100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S4682200 <- factor(data$S4682200,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S4682300 <- factor(data$S4682300,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S4812800 <- factor(data$S4812800,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$S4812900 <- factor(data$S4812900,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,97.0),
labels=c("Live with parent(s)",
"Live with spouse's/partner's parent(s)",
"Housing is part of job compensation; live-in servant; housekeeper; gardener; farm laborer",
"Housing is a gift paid for by an HU resident other than R or spouse/partner",
"Housing is a gift paid for by a friend or relative outside of the HU",
"Housing paid for by a government agency/welfare/charitable institution",
"Sold home, not moved out of it yet",
"Living in house which R will inherit; estate in progress",
"Living in temporary quarters (garage, shed) while home is under construction",
"Live here without formal arrangements; staying temporarily; squatting",
"Other"))
  data$S4919500 <- factor(data$S4919500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$S5405600 <- factor(data$S5405600,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$S5423200 <- factor(data$S5423200,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$S5436300 <- factor(data$S5436300,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$S6511400 <- factor(data$S6511400,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$S6511500 <- factor(data$S6511500,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,97.0),
labels=c("Live with parent(s)",
"Live with spouse's/partner's parent(s)",
"Housing is part of job compensation; live-in servant; housekeeper; gardener; farm laborer",
"Housing is a gift paid for by an HU resident other than R or spouse/partner",
"Housing is a gift paid for by a friend or relative outside of the HU",
"Housing paid for by a government agency/welfare/charitable institution",
"Sold home, not moved out of it yet",
"Living in house which R will inherit; estate in progress",
"Living in temporary quarters (garage, shed) while home is under construction",
"Live here without formal arrangements; staying temporarily; squatting",
"Other"))
  data$S6661100 <- factor(data$S6661100,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$S7506100 <- factor(data$S7506100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$S7509600 <- factor(data$S7509600,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$S7525300 <- factor(data$S7525300,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$S7537100 <- factor(data$S7537100,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$S8332300 <- factor(data$S8332300,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S8332400 <- factor(data$S8332400,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S8332500 <- factor(data$S8332500,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S8332600 <- factor(data$S8332600,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S8332700 <- factor(data$S8332700,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$S8509800 <- factor(data$S8509800,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$S8509900 <- factor(data$S8509900,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,97.0),
labels=c("Live with parent(s)",
"Live with spouse's/partner's parent(s)",
"Housing is part of job compensation; live-in servant; housekeeper; gardener; farm laborer",
"Housing is a gift paid for by an HU resident other than R or spouse/partner",
"Housing is a gift paid for by a friend or relative outside of the HU",
"Housing paid for by a government agency/welfare/charitable institution",
"Sold home, not moved out of it yet",
"Living in house which R will inherit; estate in progress",
"Living in temporary quarters (garage, shed) while home is under construction",
"Live here without formal arrangements; staying temporarily; squatting",
"Other"))
  data$S8644200 <- factor(data$S8644200,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$S8646800 <- factor(data$S8646800,
levels=c(1.0,2.0,3.0,4.0),
labels=c("EXCITING",
"PRETTY ROUTINE",
"DULL",
"NO OPINION"))
  data$S8647900 <- factor(data$S8647900,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("VERY GOOD TIME",
"PRETTY GOOD TIME",
"PRETTY BAD TIME",
"VERY BAD TIME",
"MIXED"))
  data$T0009400 <- factor(data$T0009400,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$T0012900 <- factor(data$T0012900,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$T0025600 <- factor(data$T0025600,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$T0033700 <- factor(data$T0033700,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$T0060700 <- factor(data$T0060700,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$T0902900 <- factor(data$T0902900,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$T0903000 <- factor(data$T0903000,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,97.0),
labels=c("Live with parent(s)",
"Live with spouse's/partner's parent(s)",
"Housing is part of job compensation; live-in servant; housekeeper; gardener; farm laborer",
"Housing is a gift paid for by an HU resident other than R or spouse/partner",
"Housing is a gift paid for by a friend or relative outside of the HU",
"Housing paid for by a government agency/welfare/charitable institution",
"Sold home, not moved out of it yet",
"Living in house which R will inherit; estate in progress",
"Living in temporary quarters (garage, shed) while home is under construction",
"Live here without formal arrangements; staying temporarily; squatting",
"Other"))
  data$T1049500 <- factor(data$T1049500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$T1049600 <- factor(data$T1049600,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat fruit",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T1049700 <- factor(data$T1049700,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat vegetables",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T1049800 <- factor(data$T1049800,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7"))
  data$T1049900 <- factor(data$T1049900,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("None",
"Less than 1 hour a week",
"1 to 3 hours a week",
"4 to 6 hours a week",
"7 to 9 hours a week",
"10 hours or more a week"))
  data$T1050000 <- factor(data$T1050000,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Less than 2 hours per week",
"3 to 10 hours a week",
"11 to 20 hours a week",
"21 to 30 hours a week",
"31 to 40 hours a week",
"More than 40 hours a week"))
  data$T1050200 <- factor(data$T1050200,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I DID NOT DRIVE IN THE PAST 30 DAYS"))
  data$T1050300 <- factor(data$T1050300,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I HAVE NOT BEEN A PASSENGER IN THE PAST 30 DAYS"))
  data$T1068300 <- factor(data$T1068300,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("VERY GOOD TIME",
"PRETTY GOOD TIME",
"PRETTY BAD TIME",
"VERY BAD TIME",
"MIXED"))
  data$T2011100 <- factor(data$T2011100,
levels=c(23.0,24.0,25.0,26.0,27.0,28.0,29.0),
labels=c("23",
"24",
"25",
"26",
"27",
"28",
"29: Age 29 or greater"))
  data$T2012100 <- factor(data$T2012100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$T2015800 <- factor(data$T2015800,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$T2020500 <- factor(data$T2020500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$T2021300 <- factor(data$T2021300,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$T2043400 <- factor(data$T2043400,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$T2782600 <- factor(data$T2782600,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T2782700 <- factor(data$T2782700,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T2782800 <- factor(data$T2782800,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T2782900 <- factor(data$T2782900,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T2783000 <- factor(data$T2783000,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T3014700 <- factor(data$T3014700,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$T3014800 <- factor(data$T3014800,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,97.0),
labels=c("LIVE WITH PARENT(s)",
"LIVE WITH SPOUSE'S/PARTNER'S PARENT(s)",
"HOUSING IS PART OF JOB COMPENSATION; LIVE-IN SERVANT; HOUSEKEEPER; GARDENER; FARM LABORER",
"HOUSING IS A GIFT PAID FOR BY AN HU RESIDENT OTHER THAN R OR SPOUSE/PARTNER",
"HOUSING IS A GIFT PAID FOR BY A FRIEND OR RELATIVE OUTSIDE OF THE HU",
"HOUSING PAID FOR BY A GOVERNMENT AGENCY/WELFARE/CHARITABLE INSTITUTION",
"SOLD HOME, NOT MOVED OUT OF IT YET",
"LIVING IN HOUSE WHICH R WILL INHERIT; ESTATE IN PROGRESS",
"LIVING IN TEMPORARY QUARTERS (GARAGE, SHED) WHILE HOME IS UNDER CONSTRUCTION",
"LIVE HERE WITHOUT FORMAL ARRANGEMENTS; STAYING TEMPORARILY; SQUATTING",
"OTHER"))
  data$T3144600 <- factor(data$T3144600,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$T3144800 <- factor(data$T3144800,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat fruit",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T3144900 <- factor(data$T3144900,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat vegetables",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T3145000 <- factor(data$T3145000,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7"))
  data$T3145100 <- factor(data$T3145100,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("None",
"Less than 1 hour a week",
"1 to 3 hours a week",
"4 to 6 hours a week",
"7 to 9 hours a week",
"10 hours or more a week"))
  data$T3145200 <- factor(data$T3145200,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Less than 2 hours per week",
"3 to 10 hours a week",
"11 to 20 hours a week",
"21 to 30 hours a week",
"31 to 40 hours a week",
"More than 40 hours a week"))
  data$T3145400 <- factor(data$T3145400,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I DID NOT DRIVE IN THE PAST 30 DAYS"))
  data$T3145500 <- factor(data$T3145500,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I HAVE NOT BEEN A PASSENGER IN THE PAST 30 DAYS"))
  data$T3162400 <- factor(data$T3162400,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("1 EXTREMELY DISSATISFIED",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 EXTREMELY SATISIFIED"))
  data$T3602100 <- factor(data$T3602100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$T3606100 <- factor(data$T3606100,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$T3611200 <- factor(data$T3611200,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$T3612000 <- factor(data$T3612000,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$T3645400 <- factor(data$T3645400,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$T4562200 <- factor(data$T4562200,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$T4562300 <- factor(data$T4562300,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat fruit",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T4562400 <- factor(data$T4562400,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat vegetables",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T4565400 <- factor(data$T4565400,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("None",
"Less than 1 hour a week",
"1 to 3 hours a week",
"4 to 6 hours a week",
"7 to 9 hours a week",
"10 hours or more a week"))
  data$T4565500 <- factor(data$T4565500,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Less than 2 hours per week",
"3 to 10 hours a week",
"11 to 20 hours a week",
"21 to 30 hours a week",
"31 to 40 hours a week",
"More than 40 hours a week"))
  data$T4565700 <- factor(data$T4565700,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I DID NOT DRIVE IN THE PAST 30 DAYS"))
  data$T4565800 <- factor(data$T4565800,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I HAVE NOT BEEN A PASSENGER IN THE PAST 30 DAYS"))
  data$T5201400 <- factor(data$T5201400,
levels=c(25.0,26.0,27.0,28.0,29.0,30.0,31.0),
labels=c("25",
"26",
"27",
"28",
"29",
"30",
"31"))
  data$T5202300 <- factor(data$T5202300,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$T5206500 <- factor(data$T5206500,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$T5211600 <- factor(data$T5211600,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$T5211900 <- factor(data$T5211900,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$T6143700 <- factor(data$T6143700,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T6143800 <- factor(data$T6143800,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T6143900 <- factor(data$T6143900,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T6144000 <- factor(data$T6144000,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T6144100 <- factor(data$T6144100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$T6206400 <- factor(data$T6206400,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$T6206500 <- factor(data$T6206500,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat fruit",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T6206600 <- factor(data$T6206600,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat vegetables",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T6209600 <- factor(data$T6209600,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("None",
"Less than 1 hour a week",
"1 to 3 hours a week",
"4 to 6 hours a week",
"7 to 9 hours a week",
"10 hours or more a week"))
  data$T6209700 <- factor(data$T6209700,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Less than 2 hours per week",
"3 to 10 hours a week",
"11 to 20 hours a week",
"21 to 30 hours a week",
"31 to 40 hours a week",
"More than 40 hours a week"))
  data$T6209900 <- factor(data$T6209900,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I DID NOT DRIVE IN THE PAST 30 DAYS"))
  data$T6210000 <- factor(data$T6210000,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I HAVE NOT BEEN A PASSENGER IN THE PAST 30 DAYS"))
  data$T6651300 <- factor(data$T6651300,
levels=c(26.0,27.0,28.0,29.0,30.0,31.0,32.0),
labels=c("26",
"27",
"28",
"29",
"30",
"31",
"32"))
  data$T6652100 <- factor(data$T6652100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$T6656300 <- factor(data$T6656300,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$T6663000 <- factor(data$T6663000,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$T6663300 <- factor(data$T6663300,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$T7703800 <- factor(data$T7703800,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$T7703900 <- factor(data$T7703900,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat fruit",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T7704000 <- factor(data$T7704000,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat vegetables",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$T7707000 <- factor(data$T7707000,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("None",
"Less than 1 hour a week",
"1 to 3 hours a week",
"4 to 6 hours a week",
"7 to 9 hours a week",
"10 hours or more a week"))
  data$T7707100 <- factor(data$T7707100,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Less than 2 hours per week",
"3 to 10 hours a week",
"11 to 20 hours a week",
"21 to 30 hours a week",
"31 to 40 hours a week",
"More than 40 hours a week"))
  data$T7707300 <- factor(data$T7707300,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I DID NOT DRIVE IN THE PAST 30 DAYS"))
  data$T7707400 <- factor(data$T7707400,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I HAVE NOT BEEN A PASSENGER IN THE PAST 30 DAYS"))
  data$T8123600 <- factor(data$T8123600,
levels=c(28.0,29.0,30.0,31.0,32.0,33.0,34.0),
labels=c("28",
"29",
"30",
"31",
"32",
"33",
"34"))
  data$T8123700 <- factor(data$T8123700,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$T8128700 <- factor(data$T8128700,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$T8134200 <- factor(data$T8134200,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$T8134500 <- factor(data$T8134500,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$T9093100 <- factor(data$T9093100,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$U0001800 <- factor(data$U0001800,
levels=c(28.0,29.0,30.0,31.0,32.0,33.0,34.0,35.0,36.0),
labels=c("28",
"29",
"30",
"31",
"32",
"33",
"34",
"35",
"36"))
  data$U0001900 <- factor(data$U0001900,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$U0008600 <- factor(data$U0008600,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Citizen, born in the U.S.",
"Citizen, naturalized",
"Applicant for Naturalization",
"Permanent Resident",
"Applicant for Residence",
"Other"))
  data$U0014700 <- factor(data$U0014700,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$U0015000 <- factor(data$U0015000,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$U1030700 <- factor(data$U1030700,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$U1030800 <- factor(data$U1030800,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$U1030900 <- factor(data$U1030900,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$U1031000 <- factor(data$U1031000,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$U1031100 <- factor(data$U1031100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("All of the time",
"Most of the time",
"Some of the time",
"None of the time"))
  data$U1096500 <- factor(data$U1096500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$U1096600 <- factor(data$U1096600,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat fruit",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$U1096700 <- factor(data$U1096700,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0),
labels=c("I do not typically eat vegetables",
"1 to 3 times",
"4 to 6 times",
"1 time per day",
"2 times per day",
"3 times per day",
"4 or more times per day"))
  data$U1099700 <- factor(data$U1099700,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("None",
"Less than 1 hour a week",
"1 to 3 hours a week",
"4 to 6 hours a week",
"7 to 9 hours a week",
"10 hours or more a week"))
  data$U1099800 <- factor(data$U1099800,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Less than 2 hours per week",
"3 to 10 hours a week",
"11 to 20 hours a week",
"21 to 30 hours a week",
"31 to 40 hours a week",
"More than 40 hours a week"))
  data$U1100000 <- factor(data$U1100000,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I DID NOT DRIVE IN THE PAST 30 DAYS"))
  data$U1100100 <- factor(data$U1100100,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I HAVE NOT BEEN A PASSENGER IN THE PAST 30 DAYS"))
  data$U1838500 <- factor(data$U1838500,
levels=c(32.0,33.0,34.0,35.0,36.0,37.0,38.0,39.0,40.0),
labels=c("32",
"33",
"34",
"35",
"36",
"37",
"38",
"39",
"40"))
  data$U1838600 <- factor(data$U1838600,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$U1852500 <- factor(data$U1852500,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$U1853200 <- factor(data$U1853200,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$U2962600 <- factor(data$U2962600,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$U2962800 <- factor(data$U2962800,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("None",
"Less than 1 hour a week",
"1 to 3 hours a week",
"4 to 6 hours a week",
"7 to 9 hours a week",
"10 hours or more a week"))
  data$U2962900 <- factor(data$U2962900,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("Less than 2 hours per week",
"3 to 10 hours a week",
"11 to 20 hours a week",
"21 to 30 hours a week",
"31 to 40 hours a week",
"More than 40 hours a week"))
  data$U2963200 <- factor(data$U2963200,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I DID NOT DRIVE IN THE PAST 30 DAYS"))
  data$U2963300 <- factor(data$U2963300,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("0 times",
"1 time",
"2 or 3 times",
"4 or 5 times",
"6 or more times",
"I HAVE NOT BEEN A PASSENGER IN THE PAST 30 DAYS"))
  data$U2964800 <- factor(data$U2964800,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U2964900 <- factor(data$U2964900,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U2965000 <- factor(data$U2965000,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U2965100 <- factor(data$U2965100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U2965200 <- factor(data$U2965200,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U3437900 <- factor(data$U3437900,
levels=c(33.0,34.0,35.0,36.0,37.0,38.0,39.0,40.0),
labels=c("33",
"34",
"35",
"36",
"37",
"38",
"39",
"40"))
  data$U3438000 <- factor(data$U3438000,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$U3451600 <- factor(data$U3451600,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$U3453600 <- factor(data$U3453600,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$U4368000 <- factor(data$U4368000,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$U4369900 <- factor(data$U4369900,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U4370000 <- factor(data$U4370000,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U4370100 <- factor(data$U4370100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U4370200 <- factor(data$U4370200,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U4370300 <- factor(data$U4370300,
levels=c(1.0,2.0,3.0,4.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"SOME OF THE TIME",
"NONE OF THE TIME"))
  data$U4943000 <- factor(data$U4943000,
levels=c(36.0,37.0,38.0,39.0,40.0,41.0,42.0),
labels=c("36",
"37",
"38",
"39",
"40",
"41",
"42"))
  data$U4943100 <- factor(data$U4943100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("Northeast (CT, ME, MA, NH, NJ, NY, PA, RI, VT)",
"North Central (IL, IN, IA, KS, MI, MN, MO, NE, OH, ND, SD, WI)",
"South (AL, AR, DE, DC, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN , TX, VA, WV)",
"West (AK, AZ, CA, CO, HI, ID, MT, NV, NM, OR, UT, WA, WY)"))
  data$U4954700 <- factor(data$U4954700,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Not in CBSA",
"In CBSA, not in central city",
"In CBSA, in central city",
"In CBSA, not known",
"Not in country"))
  data$U4956900 <- factor(data$U4956900,
levels=c(0.0,1.0,2.0),
labels=c("Rural",
"Urban",
"Unknown"))
  data$U5860700 <- factor(data$U5860700,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("Excellent",
"Very good",
"Good",
"Fair",
"Poor"))
  data$U5867900 <- factor(data$U5867900,
levels=c(0.0,1.0),
labels=c("NO",
"YES"))
  data$U5868000 <- factor(data$U5868000,
levels=c(0.0,1.0),
labels=c("NOT SELECTED",
"A cellular plan for a smartphone or other mobile device"))
  data$U5868100 <- factor(data$U5868100,
levels=c(1.0,2.0,3.0,4.0),
labels=c("MOSTLY ON CELLPHONE",
"MOSTLY ON SOMETHING ELSE",
"BOTH EQUALLY",
"DON'T USE THE INTERNET"))
  data$Z0397600 <- factor(data$Z0397600,
levels=c(1.0,2.0,3.0,4.0,5.0),
labels=c("RESPONDENT OWNS",
"RESPONDENT AND SPOUSE/PARTNER JOINTLY OWN",
"SPOUSE/PARTNER OWNS SEPARATELY FROM RESPONDENT",
"RENTS",
"OTHER, NEITHER OWNS NOR RENTS"))
  data$Z0397700 <- factor(data$Z0397700,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,97.0),
labels=c("LIVE WITH PARENT(s)",
"LIVE WITH SPOUSE'S/PARTNER'S PARENT(s)",
"HOUSING IS PART OF JOB COMPENSATION; LIVE-IN SERVANT; HOUSEKEEPER; GARDENER; FARM LABORER",
"HOUSING IS A GIFT PAID FOR BY AN HU RESIDENT OTHER THAN R OR SPOUSE/PARTNER",
"HOUSING IS A GIFT PAID FOR BY A FRIEND OR RELATIVE OUTSIDE OF THE HU",
"HOUSING PAID FOR BY A GOVERNMENT AGENCY/WELFARE/CHARITABLE INSTITUTION",
"SOLD HOME, NOT MOVED OUT OF IT YET",
"LIVING IN HOUSE WHICH R WILL INHERIT; ESTATE IN PROGRESS",
"LIVING IN TEMPORARY QUARTERS (GARAGE, SHED) WHILE HOME IS UNDER CONSTRUCTION",
"LIVE HERE WITHOUT FORMAL ARRANGEMENTS; STAYING TEMPORARILY; SQUATTING",
"OTHER"))
  data$Z0530200 <- factor(data$Z0530200,
levels=c(1.0,2.0,3.0,4.0,5.0,6.0),
labels=c("ALL OF THE TIME",
"MOST OF THE TIME",
"A GOOD BIT OF THE TIME",
"SOME OF THE TIME",
"A LITTLE OF THE TIME",
"NONE OF THE TIME"))
return(data)
}


# If there are values not categorized they will be represented as NA

vallabels_continuous = function(data) {
data$R0000100[1.0 <= data$R0000100 & data$R0000100 <= 999.0] <- 1.0
data$R0000100[1000.0 <= data$R0000100 & data$R0000100 <= 1999.0] <- 1000.0
data$R0000100[2000.0 <= data$R0000100 & data$R0000100 <= 2999.0] <- 2000.0
data$R0000100[3000.0 <= data$R0000100 & data$R0000100 <= 3999.0] <- 3000.0
data$R0000100[4000.0 <= data$R0000100 & data$R0000100 <= 4999.0] <- 4000.0
data$R0000100[5000.0 <= data$R0000100 & data$R0000100 <= 5999.0] <- 5000.0
data$R0000100[6000.0 <= data$R0000100 & data$R0000100 <= 6999.0] <- 6000.0
data$R0000100[7000.0 <= data$R0000100 & data$R0000100 <= 7999.0] <- 7000.0
data$R0000100[8000.0 <= data$R0000100 & data$R0000100 <= 8999.0] <- 8000.0
data$R0000100[9000.0 <= data$R0000100 & data$R0000100 <= 9999.0] <- 9000.0
data$R0000100 <- factor(data$R0000100,
levels=c(0.0,1.0,1000.0,2000.0,3000.0,4000.0,5000.0,6000.0,7000.0,8000.0,9000.0),
labels=c("0",
"1 TO 999",
"1000 TO 1999",
"2000 TO 2999",
"3000 TO 3999",
"4000 TO 4999",
"5000 TO 5999",
"6000 TO 6999",
"7000 TO 7999",
"8000 TO 8999",
"9000 TO 9999"))
data$R0321200[1.0 <= data$R0321200 & data$R0321200 <= 10.0] <- 1.0
data$R0321200[11.0 <= data$R0321200 & data$R0321200 <= 20.0] <- 11.0
data$R0321200[21.0 <= data$R0321200 & data$R0321200 <= 30.0] <- 21.0
data$R0321200[31.0 <= data$R0321200 & data$R0321200 <= 40.0] <- 31.0
data$R0321200[41.0 <= data$R0321200 & data$R0321200 <= 50.0] <- 41.0
data$R0321200[51.0 <= data$R0321200 & data$R0321200 <= 60.0] <- 51.0
data$R0321200[61.0 <= data$R0321200 & data$R0321200 <= 70.0] <- 61.0
data$R0321200[71.0 <= data$R0321200 & data$R0321200 <= 80.0] <- 71.0
data$R0321200[81.0 <= data$R0321200 & data$R0321200 <= 90.0] <- 81.0
data$R0321200[91.0 <= data$R0321200 & data$R0321200 <= 100.0] <- 91.0
data$R0321200 <- factor(data$R0321200,
levels=c(0.0,1.0,11.0,21.0,31.0,41.0,51.0,61.0,71.0,81.0,91.0),
labels=c("0: 0%",
"1 TO 10: 1%-10%",
"11 TO 20: 11%-20%",
"21 TO 30: 21%-30%",
"31 TO 40: 31%-40%",
"41 TO 50: 41%-50%",
"51 TO 60: 51%-60%",
"61 TO 70: 61%-70%",
"71 TO 80: 71%-80%",
"81 TO 90: 81%-90%",
"91 TO 100: 91%-100%"))
data$R0536600[0.0 <= data$R0536600 & data$R0536600 <= 11.0] <- 0.0
data$R0536600[19.0 <= data$R0536600 & data$R0536600 <= 999.0] <- 19.0
data$R0536600 <- factor(data$R0536600,
levels=c(0.0,12.0,13.0,14.0,15.0,16.0,17.0,18.0,19.0),
labels=c("0 TO 11: LESS THAN 12",
"12",
"13",
"14",
"15",
"16",
"17",
"18",
"19 TO 999: GREATER THAN 18"))
data$R1193900[0.0 <= data$R1193900 & data$R1193900 <= 139.0] <- 0.0
data$R1193900[140.0 <= data$R1193900 & data$R1193900 <= 150.0] <- 140.0
data$R1193900[151.0 <= data$R1193900 & data$R1193900 <= 160.0] <- 151.0
data$R1193900[161.0 <= data$R1193900 & data$R1193900 <= 170.0] <- 161.0
data$R1193900[171.0 <= data$R1193900 & data$R1193900 <= 180.0] <- 171.0
data$R1193900[181.0 <= data$R1193900 & data$R1193900 <= 190.0] <- 181.0
data$R1193900[191.0 <= data$R1193900 & data$R1193900 <= 200.0] <- 191.0
data$R1193900[201.0 <= data$R1193900 & data$R1193900 <= 210.0] <- 201.0
data$R1193900[211.0 <= data$R1193900 & data$R1193900 <= 220.0] <- 211.0
data$R1193900[221.0 <= data$R1193900 & data$R1193900 <= 230.0] <- 221.0
data$R1193900 <- factor(data$R1193900,
levels=c(0.0,140.0,151.0,161.0,171.0,181.0,191.0,201.0,211.0,221.0),
labels=c("0 TO 139: <140",
"140 TO 150",
"151 TO 160",
"161 TO 170",
"171 TO 180",
"181 TO 190",
"191 TO 200",
"201 TO 210",
"211 TO 220",
"221 TO 230"))
data$R1194000[0.0 <= data$R1194000 & data$R1194000 <= 11.0] <- 0.0
data$R1194000[19.0 <= data$R1194000 & data$R1194000 <= 999.0] <- 19.0
data$R1194000 <- factor(data$R1194000,
levels=c(0.0,12.0,13.0,14.0,15.0,16.0,17.0,18.0,19.0),
labels=c("0 TO 11: LESS THAN 12",
"12",
"13",
"14",
"15",
"16",
"17",
"18",
"19 TO 999: GREATER THAN 18"))
data$R1194100[0.0 <= data$R1194100 & data$R1194100 <= 11.0] <- 0.0
data$R1194100[19.0 <= data$R1194100 & data$R1194100 <= 999.0] <- 19.0
data$R1194100 <- factor(data$R1194100,
levels=c(0.0,12.0,13.0,14.0,15.0,16.0,17.0,18.0,19.0),
labels=c("0 TO 11: LESS THAN 12",
"12",
"13",
"14",
"15",
"16",
"17",
"18",
"19 TO 999: GREATER THAN 18"))
data$R1200100[1.0 <= data$R1200100 & data$R1200100 <= 4.0] <- 1.0
data$R1200100[5.0 <= data$R1200100 & data$R1200100 <= 8.0] <- 5.0
data$R1200100[9.0 <= data$R1200100 & data$R1200100 <= 14.0] <- 9.0
data$R1200100[15.0 <= data$R1200100 & data$R1200100 <= 19.0] <- 15.0
data$R1200100[20.0 <= data$R1200100 & data$R1200100 <= 24.0] <- 20.0
data$R1200100[25.0 <= data$R1200100 & data$R1200100 <= 29.0] <- 25.0
data$R1200100[30.0 <= data$R1200100 & data$R1200100 <= 34.0] <- 30.0
data$R1200100[35.0 <= data$R1200100 & data$R1200100 <= 39.0] <- 35.0
data$R1200100[40.0 <= data$R1200100 & data$R1200100 <= 44.0] <- 40.0
data$R1200100[45.0 <= data$R1200100 & data$R1200100 <= 49.0] <- 45.0
data$R1200100[50.0 <= data$R1200100 & data$R1200100 <= 9.9999999E7] <- 50.0
data$R1200100 <- factor(data$R1200100,
levels=c(0.0,1.0,5.0,9.0,15.0,20.0,25.0,30.0,35.0,40.0,45.0,50.0),
labels=c("0",
"1 TO 4",
"5 TO 8",
"9 TO 14",
"15 TO 19",
"20 TO 24",
"25 TO 29",
"30 TO 34",
"35 TO 39",
"40 TO 44",
"45 TO 49",
"50 TO 99999999: 50+"))
data$R1200200[1.0 <= data$R1200200 & data$R1200200 <= 4.0] <- 1.0
data$R1200200[5.0 <= data$R1200200 & data$R1200200 <= 8.0] <- 5.0
data$R1200200[9.0 <= data$R1200200 & data$R1200200 <= 14.0] <- 9.0
data$R1200200[15.0 <= data$R1200200 & data$R1200200 <= 19.0] <- 15.0
data$R1200200[20.0 <= data$R1200200 & data$R1200200 <= 24.0] <- 20.0
data$R1200200[25.0 <= data$R1200200 & data$R1200200 <= 29.0] <- 25.0
data$R1200200[30.0 <= data$R1200200 & data$R1200200 <= 34.0] <- 30.0
data$R1200200[35.0 <= data$R1200200 & data$R1200200 <= 39.0] <- 35.0
data$R1200200[40.0 <= data$R1200200 & data$R1200200 <= 44.0] <- 40.0
data$R1200200[45.0 <= data$R1200200 & data$R1200200 <= 49.0] <- 45.0
data$R1200200[50.0 <= data$R1200200 & data$R1200200 <= 9.9999999E7] <- 50.0
data$R1200200 <- factor(data$R1200200,
levels=c(0.0,1.0,5.0,9.0,15.0,20.0,25.0,30.0,35.0,40.0,45.0,50.0),
labels=c("0",
"1 TO 4",
"5 TO 8",
"9 TO 14",
"15 TO 19",
"20 TO 24",
"25 TO 29",
"30 TO 34",
"35 TO 39",
"40 TO 44",
"45 TO 49",
"50 TO 99999999: 50+"))
data$R2553400[0.0 <= data$R2553400 & data$R2553400 <= 139.0] <- 0.0
data$R2553400[140.0 <= data$R2553400 & data$R2553400 <= 150.0] <- 140.0
data$R2553400[151.0 <= data$R2553400 & data$R2553400 <= 160.0] <- 151.0
data$R2553400[161.0 <= data$R2553400 & data$R2553400 <= 170.0] <- 161.0
data$R2553400[171.0 <= data$R2553400 & data$R2553400 <= 180.0] <- 171.0
data$R2553400[181.0 <= data$R2553400 & data$R2553400 <= 190.0] <- 181.0
data$R2553400[191.0 <= data$R2553400 & data$R2553400 <= 200.0] <- 191.0
data$R2553400[201.0 <= data$R2553400 & data$R2553400 <= 210.0] <- 201.0
data$R2553400[211.0 <= data$R2553400 & data$R2553400 <= 220.0] <- 211.0
data$R2553400[221.0 <= data$R2553400 & data$R2553400 <= 230.0] <- 221.0
data$R2553400[231.0 <= data$R2553400 & data$R2553400 <= 240.0] <- 231.0
data$R2553400 <- factor(data$R2553400,
levels=c(0.0,140.0,151.0,161.0,171.0,181.0,191.0,201.0,211.0,221.0,231.0),
labels=c("0 TO 139: <140",
"140 TO 150",
"151 TO 160",
"161 TO 170",
"171 TO 180",
"181 TO 190",
"191 TO 200",
"201 TO 210",
"211 TO 220",
"221 TO 230",
"231 TO 240"))
data$R2553500[0.0 <= data$R2553500 & data$R2553500 <= 11.0] <- 0.0
data$R2553500[19.0 <= data$R2553500 & data$R2553500 <= 999.0] <- 19.0
data$R2553500 <- factor(data$R2553500,
levels=c(0.0,12.0,13.0,14.0,15.0,16.0,17.0,18.0,19.0),
labels=c("0 TO 11: LESS THAN 12",
"12",
"13",
"14",
"15",
"16",
"17",
"18",
"19 TO 999: GREATER THAN 18"))
data$R3876200[0.0 <= data$R3876200 & data$R3876200 <= 139.0] <- 0.0
data$R3876200[140.0 <= data$R3876200 & data$R3876200 <= 150.0] <- 140.0
data$R3876200[151.0 <= data$R3876200 & data$R3876200 <= 160.0] <- 151.0
data$R3876200[161.0 <= data$R3876200 & data$R3876200 <= 170.0] <- 161.0
data$R3876200[171.0 <= data$R3876200 & data$R3876200 <= 180.0] <- 171.0
data$R3876200[181.0 <= data$R3876200 & data$R3876200 <= 190.0] <- 181.0
data$R3876200[191.0 <= data$R3876200 & data$R3876200 <= 200.0] <- 191.0
data$R3876200[201.0 <= data$R3876200 & data$R3876200 <= 210.0] <- 201.0
data$R3876200[211.0 <= data$R3876200 & data$R3876200 <= 220.0] <- 211.0
data$R3876200[221.0 <= data$R3876200 & data$R3876200 <= 230.0] <- 221.0
data$R3876200[231.0 <= data$R3876200 & data$R3876200 <= 240.0] <- 231.0
data$R3876200[241.0 <= data$R3876200 & data$R3876200 <= 250.0] <- 241.0
data$R3876200 <- factor(data$R3876200,
levels=c(0.0,140.0,151.0,161.0,171.0,181.0,191.0,201.0,211.0,221.0,231.0,241.0),
labels=c("0 TO 139: <140",
"140 TO 150",
"151 TO 160",
"161 TO 170",
"171 TO 180",
"181 TO 190",
"191 TO 200",
"201 TO 210",
"211 TO 220",
"221 TO 230",
"231 TO 240",
"241 TO 250"))
data$R3876300[0.0 <= data$R3876300 & data$R3876300 <= 11.0] <- 0.0
data$R3876300[19.0 <= data$R3876300 & data$R3876300 <= 999.0] <- 19.0
data$R3876300 <- factor(data$R3876300,
levels=c(0.0,12.0,13.0,14.0,15.0,16.0,17.0,18.0,19.0),
labels=c("0 TO 11: LESS THAN 12",
"12",
"13",
"14",
"15",
"16",
"17",
"18",
"19 TO 999: GREATER THAN 18"))
data$R5453600[0.0 <= data$R5453600 & data$R5453600 <= 139.0] <- 0.0
data$R5453600[140.0 <= data$R5453600 & data$R5453600 <= 150.0] <- 140.0
data$R5453600[151.0 <= data$R5453600 & data$R5453600 <= 160.0] <- 151.0
data$R5453600[161.0 <= data$R5453600 & data$R5453600 <= 170.0] <- 161.0
data$R5453600[171.0 <= data$R5453600 & data$R5453600 <= 180.0] <- 171.0
data$R5453600[181.0 <= data$R5453600 & data$R5453600 <= 190.0] <- 181.0
data$R5453600[191.0 <= data$R5453600 & data$R5453600 <= 200.0] <- 191.0
data$R5453600[201.0 <= data$R5453600 & data$R5453600 <= 210.0] <- 201.0
data$R5453600[211.0 <= data$R5453600 & data$R5453600 <= 220.0] <- 211.0
data$R5453600[221.0 <= data$R5453600 & data$R5453600 <= 230.0] <- 221.0
data$R5453600[231.0 <= data$R5453600 & data$R5453600 <= 240.0] <- 231.0
data$R5453600[241.0 <= data$R5453600 & data$R5453600 <= 250.0] <- 241.0
data$R5453600[251.0 <= data$R5453600 & data$R5453600 <= 260.0] <- 251.0
data$R5453600 <- factor(data$R5453600,
levels=c(0.0,140.0,151.0,161.0,171.0,181.0,191.0,201.0,211.0,221.0,231.0,241.0,251.0),
labels=c("0 TO 139: <140",
"140 TO 150",
"151 TO 160",
"161 TO 170",
"171 TO 180",
"181 TO 190",
"191 TO 200",
"201 TO 210",
"211 TO 220",
"221 TO 230",
"231 TO 240",
"241 TO 250",
"251 TO 260"))
data$R5453700[21.0 <= data$R5453700 & data$R5453700 <= 100.0] <- 21.0
data$R5453700 <- factor(data$R5453700,
levels=c(14.0,15.0,16.0,17.0,18.0,19.0,20.0,21.0),
labels=c("14",
"15",
"16",
"17",
"18",
"19",
"20",
"21 TO 100: Age 21 and older"))
data$R7215900[0.0 <= data$R7215900 & data$R7215900 <= 139.0] <- 0.0
data$R7215900[140.0 <= data$R7215900 & data$R7215900 <= 150.0] <- 140.0
data$R7215900[151.0 <= data$R7215900 & data$R7215900 <= 160.0] <- 151.0
data$R7215900[161.0 <= data$R7215900 & data$R7215900 <= 170.0] <- 161.0
data$R7215900[171.0 <= data$R7215900 & data$R7215900 <= 180.0] <- 171.0
data$R7215900[181.0 <= data$R7215900 & data$R7215900 <= 190.0] <- 181.0
data$R7215900[191.0 <= data$R7215900 & data$R7215900 <= 200.0] <- 191.0
data$R7215900[201.0 <= data$R7215900 & data$R7215900 <= 210.0] <- 201.0
data$R7215900[211.0 <= data$R7215900 & data$R7215900 <= 220.0] <- 211.0
data$R7215900[221.0 <= data$R7215900 & data$R7215900 <= 230.0] <- 221.0
data$R7215900[231.0 <= data$R7215900 & data$R7215900 <= 240.0] <- 231.0
data$R7215900[241.0 <= data$R7215900 & data$R7215900 <= 250.0] <- 241.0
data$R7215900[251.0 <= data$R7215900 & data$R7215900 <= 260.0] <- 251.0
data$R7215900[261.0 <= data$R7215900 & data$R7215900 <= 270.0] <- 261.0
data$R7215900 <- factor(data$R7215900,
levels=c(0.0,140.0,151.0,161.0,171.0,181.0,191.0,201.0,211.0,221.0,231.0,241.0,251.0,261.0),
labels=c("0 TO 139: <140",
"140 TO 150",
"151 TO 160",
"161 TO 170",
"171 TO 180",
"181 TO 190",
"191 TO 200",
"201 TO 210",
"211 TO 220",
"221 TO 230",
"231 TO 240",
"241 TO 250",
"251 TO 260",
"261 TO 270"))
data$R7216000[22.0 <= data$R7216000 & data$R7216000 <= 100.0] <- 22.0
data$R7216000 <- factor(data$R7216000,
levels=c(14.0,15.0,16.0,17.0,18.0,19.0,20.0,21.0,22.0),
labels=c("14",
"15",
"16",
"17",
"18",
"19",
"20",
"21",
"22 TO 100: Age 22 or greater"))
data$S1225300[10.0 <= data$S1225300 & data$S1225300 <= 999.0] <- 10.0
data$S1225300 <- factor(data$S1225300,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 TO 999: 10+"))
data$S1225600[10.0 <= data$S1225600 & data$S1225600 <= 999.0] <- 10.0
data$S1225600 <- factor(data$S1225600,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 TO 999: 10+"))
data$S1531300[0.0 <= data$S1531300 & data$S1531300 <= 139.0] <- 0.0
data$S1531300[140.0 <= data$S1531300 & data$S1531300 <= 150.0] <- 140.0
data$S1531300[151.0 <= data$S1531300 & data$S1531300 <= 160.0] <- 151.0
data$S1531300[161.0 <= data$S1531300 & data$S1531300 <= 170.0] <- 161.0
data$S1531300[171.0 <= data$S1531300 & data$S1531300 <= 180.0] <- 171.0
data$S1531300[181.0 <= data$S1531300 & data$S1531300 <= 190.0] <- 181.0
data$S1531300[191.0 <= data$S1531300 & data$S1531300 <= 200.0] <- 191.0
data$S1531300[201.0 <= data$S1531300 & data$S1531300 <= 210.0] <- 201.0
data$S1531300[211.0 <= data$S1531300 & data$S1531300 <= 220.0] <- 211.0
data$S1531300[221.0 <= data$S1531300 & data$S1531300 <= 230.0] <- 221.0
data$S1531300[231.0 <= data$S1531300 & data$S1531300 <= 240.0] <- 231.0
data$S1531300[241.0 <= data$S1531300 & data$S1531300 <= 250.0] <- 241.0
data$S1531300[251.0 <= data$S1531300 & data$S1531300 <= 260.0] <- 251.0
data$S1531300[261.0 <= data$S1531300 & data$S1531300 <= 270.0] <- 261.0
data$S1531300[271.0 <= data$S1531300 & data$S1531300 <= 280.0] <- 271.0
data$S1531300[281.0 <= data$S1531300 & data$S1531300 <= 290.0] <- 281.0
data$S1531300[291.0 <= data$S1531300 & data$S1531300 <= 300.0] <- 291.0
data$S1531300 <- factor(data$S1531300,
levels=c(0.0,140.0,151.0,161.0,171.0,181.0,191.0,201.0,211.0,221.0,231.0,241.0,251.0,261.0,271.0,281.0,291.0),
labels=c("0 TO 139: <140",
"140 TO 150",
"151 TO 160",
"161 TO 170",
"171 TO 180",
"181 TO 190",
"191 TO 200",
"201 TO 210",
"211 TO 220",
"221 TO 230",
"231 TO 240",
"241 TO 250",
"251 TO 260",
"261 TO 270",
"271 TO 280",
"281 TO 290",
"291 TO 300"))
data$S1531400[22.0 <= data$S1531400 & data$S1531400 <= 100.0] <- 22.0
data$S1531400 <- factor(data$S1531400,
levels=c(14.0,15.0,16.0,17.0,18.0,19.0,20.0,21.0,22.0),
labels=c("14",
"15",
"16",
"17",
"18",
"19",
"20",
"21",
"22 TO 100: Age 22 or greater"))
data$S2000900[0.0 <= data$S2000900 & data$S2000900 <= 200.0] <- 0.0
data$S2000900[201.0 <= data$S2000900 & data$S2000900 <= 210.0] <- 201.0
data$S2000900[211.0 <= data$S2000900 & data$S2000900 <= 220.0] <- 211.0
data$S2000900[221.0 <= data$S2000900 & data$S2000900 <= 230.0] <- 221.0
data$S2000900[231.0 <= data$S2000900 & data$S2000900 <= 240.0] <- 231.0
data$S2000900[241.0 <= data$S2000900 & data$S2000900 <= 250.0] <- 241.0
data$S2000900[251.0 <= data$S2000900 & data$S2000900 <= 260.0] <- 251.0
data$S2000900[261.0 <= data$S2000900 & data$S2000900 <= 270.0] <- 261.0
data$S2000900[271.0 <= data$S2000900 & data$S2000900 <= 280.0] <- 271.0
data$S2000900[281.0 <= data$S2000900 & data$S2000900 <= 290.0] <- 281.0
data$S2000900[291.0 <= data$S2000900 & data$S2000900 <= 300.0] <- 291.0
data$S2000900 <- factor(data$S2000900,
levels=c(0.0,201.0,211.0,221.0,231.0,241.0,251.0,261.0,271.0,281.0,291.0),
labels=c("0 TO 200",
"201 TO 210",
"211 TO 220",
"221 TO 230",
"231 TO 240",
"241 TO 250",
"251 TO 260",
"261 TO 270",
"271 TO 280",
"281 TO 290",
"291 TO 300"))
data$S2001000[24.0 <= data$S2001000 & data$S2001000 <= 100.0] <- 24.0
data$S2001000 <- factor(data$S2001000,
levels=c(18.0,19.0,20.0,21.0,22.0,23.0,24.0),
labels=c("18",
"19",
"20",
"21",
"22",
"23",
"24 TO 100: Age 24 or greater"))
data$S3801000[0.0 <= data$S3801000 & data$S3801000 <= 200.0] <- 0.0
data$S3801000[201.0 <= data$S3801000 & data$S3801000 <= 220.0] <- 201.0
data$S3801000[221.0 <= data$S3801000 & data$S3801000 <= 240.0] <- 221.0
data$S3801000[241.0 <= data$S3801000 & data$S3801000 <= 260.0] <- 241.0
data$S3801000[261.0 <= data$S3801000 & data$S3801000 <= 280.0] <- 261.0
data$S3801000[281.0 <= data$S3801000 & data$S3801000 <= 300.0] <- 281.0
data$S3801000[301.0 <= data$S3801000 & data$S3801000 <= 320.0] <- 301.0
data$S3801000[321.0 <= data$S3801000 & data$S3801000 <= 340.0] <- 321.0
data$S3801000 <- factor(data$S3801000,
levels=c(0.0,201.0,221.0,241.0,261.0,281.0,301.0,321.0),
labels=c("0 TO 200: <200",
"201 TO 220",
"221 TO 240",
"241 TO 260",
"261 TO 280",
"281 TO 300",
"301 TO 320",
"321 TO 340"))
data$S3801100[24.0 <= data$S3801100 & data$S3801100 <= 100.0] <- 24.0
data$S3801100 <- factor(data$S3801100,
levels=c(18.0,19.0,20.0,21.0,22.0,23.0,24.0),
labels=c("18",
"19",
"20",
"21",
"22",
"23",
"24 TO 100: Age 24 or greater"))
data$S5400900[0.0 <= data$S5400900 & data$S5400900 <= 200.0] <- 0.0
data$S5400900[201.0 <= data$S5400900 & data$S5400900 <= 220.0] <- 201.0
data$S5400900[221.0 <= data$S5400900 & data$S5400900 <= 240.0] <- 221.0
data$S5400900[241.0 <= data$S5400900 & data$S5400900 <= 260.0] <- 241.0
data$S5400900[261.0 <= data$S5400900 & data$S5400900 <= 280.0] <- 261.0
data$S5400900[281.0 <= data$S5400900 & data$S5400900 <= 300.0] <- 281.0
data$S5400900[301.0 <= data$S5400900 & data$S5400900 <= 320.0] <- 301.0
data$S5400900[321.0 <= data$S5400900 & data$S5400900 <= 340.0] <- 321.0
data$S5400900 <- factor(data$S5400900,
levels=c(0.0,201.0,221.0,241.0,261.0,281.0,301.0,321.0),
labels=c("0 TO 200: <200",
"201 TO 220",
"221 TO 240",
"241 TO 260",
"261 TO 280",
"281 TO 300",
"301 TO 320",
"321 TO 340"))
data$S5401000[27.0 <= data$S5401000 & data$S5401000 <= 99.0] <- 27.0
data$S5401000 <- factor(data$S5401000,
levels=c(20.0,21.0,22.0,23.0,24.0,25.0,26.0,27.0),
labels=c("20",
"21",
"22",
"23",
"24",
"25",
"26",
"27 TO 99: Age 27 or greater"))
data$S7501100[0.0 <= data$S7501100 & data$S7501100 <= 200.0] <- 0.0
data$S7501100[201.0 <= data$S7501100 & data$S7501100 <= 220.0] <- 201.0
data$S7501100[221.0 <= data$S7501100 & data$S7501100 <= 240.0] <- 221.0
data$S7501100[241.0 <= data$S7501100 & data$S7501100 <= 260.0] <- 241.0
data$S7501100[261.0 <= data$S7501100 & data$S7501100 <= 280.0] <- 261.0
data$S7501100[281.0 <= data$S7501100 & data$S7501100 <= 300.0] <- 281.0
data$S7501100[301.0 <= data$S7501100 & data$S7501100 <= 320.0] <- 301.0
data$S7501100[321.0 <= data$S7501100 & data$S7501100 <= 340.0] <- 321.0
data$S7501100 <- factor(data$S7501100,
levels=c(0.0,201.0,221.0,241.0,261.0,281.0,301.0,321.0),
labels=c("0 TO 200: <200",
"201 TO 220",
"221 TO 240",
"241 TO 260",
"261 TO 280",
"281 TO 300",
"301 TO 320",
"321 TO 340"))
data$S7501200[27.0 <= data$S7501200 & data$S7501200 <= 99.0] <- 27.0
data$S7501200 <- factor(data$S7501200,
levels=c(20.0,21.0,22.0,23.0,24.0,25.0,26.0,27.0),
labels=c("20",
"21",
"22",
"23",
"24",
"25",
"26",
"27 TO 99: Age 27 or greater"))
data$T0008400[0.0 <= data$T0008400 & data$T0008400 <= 200.0] <- 0.0
data$T0008400[201.0 <= data$T0008400 & data$T0008400 <= 220.0] <- 201.0
data$T0008400[221.0 <= data$T0008400 & data$T0008400 <= 240.0] <- 221.0
data$T0008400[241.0 <= data$T0008400 & data$T0008400 <= 260.0] <- 241.0
data$T0008400[261.0 <= data$T0008400 & data$T0008400 <= 280.0] <- 261.0
data$T0008400[281.0 <= data$T0008400 & data$T0008400 <= 300.0] <- 281.0
data$T0008400[301.0 <= data$T0008400 & data$T0008400 <= 320.0] <- 301.0
data$T0008400[321.0 <= data$T0008400 & data$T0008400 <= 340.0] <- 321.0
data$T0008400[341.0 <= data$T0008400 & data$T0008400 <= 360.0] <- 341.0
data$T0008400 <- factor(data$T0008400,
levels=c(0.0,201.0,221.0,241.0,261.0,281.0,301.0,321.0,341.0),
labels=c("0 TO 200: <200",
"201 TO 220",
"221 TO 240",
"241 TO 260",
"261 TO 280",
"281 TO 300",
"301 TO 320",
"321 TO 340",
"341 TO 360"))
data$T0008500[27.0 <= data$T0008500 & data$T0008500 <= 99.0] <- 27.0
data$T0008500 <- factor(data$T0008500,
levels=c(20.0,21.0,22.0,23.0,24.0,25.0,26.0,27.0),
labels=c("20",
"21",
"22",
"23",
"24",
"25",
"26",
"27 TO 99: Age 27 or greater"))
data$T1050100[10.0 <= data$T1050100 & data$T1050100 <= 999.0] <- 10.0
data$T1050100 <- factor(data$T1050100,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 TO 999: 10+"))
data$T2011000[0.0 <= data$T2011000 & data$T2011000 <= 200.0] <- 0.0
data$T2011000[201.0 <= data$T2011000 & data$T2011000 <= 220.0] <- 201.0
data$T2011000[221.0 <= data$T2011000 & data$T2011000 <= 240.0] <- 221.0
data$T2011000[241.0 <= data$T2011000 & data$T2011000 <= 260.0] <- 241.0
data$T2011000[261.0 <= data$T2011000 & data$T2011000 <= 280.0] <- 261.0
data$T2011000[281.0 <= data$T2011000 & data$T2011000 <= 300.0] <- 281.0
data$T2011000[301.0 <= data$T2011000 & data$T2011000 <= 320.0] <- 301.0
data$T2011000[321.0 <= data$T2011000 & data$T2011000 <= 340.0] <- 321.0
data$T2011000[341.0 <= data$T2011000 & data$T2011000 <= 360.0] <- 341.0
data$T2011000 <- factor(data$T2011000,
levels=c(0.0,201.0,221.0,241.0,261.0,281.0,301.0,321.0,341.0),
labels=c("0 TO 200: <200",
"201 TO 220",
"221 TO 240",
"241 TO 260",
"261 TO 280",
"281 TO 300",
"301 TO 320",
"321 TO 340",
"341 TO 360"))
data$T3145300[10.0 <= data$T3145300 & data$T3145300 <= 999.0] <- 10.0
data$T3145300 <- factor(data$T3145300,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 TO 999: 10+"))
data$T3601400[0.0 <= data$T3601400 & data$T3601400 <= 200.0] <- 0.0
data$T3601400[201.0 <= data$T3601400 & data$T3601400 <= 220.0] <- 201.0
data$T3601400[221.0 <= data$T3601400 & data$T3601400 <= 240.0] <- 221.0
data$T3601400[241.0 <= data$T3601400 & data$T3601400 <= 260.0] <- 241.0
data$T3601400[261.0 <= data$T3601400 & data$T3601400 <= 280.0] <- 261.0
data$T3601400[281.0 <= data$T3601400 & data$T3601400 <= 300.0] <- 281.0
data$T3601400[301.0 <= data$T3601400 & data$T3601400 <= 320.0] <- 301.0
data$T3601400[321.0 <= data$T3601400 & data$T3601400 <= 340.0] <- 321.0
data$T3601400[341.0 <= data$T3601400 & data$T3601400 <= 360.0] <- 341.0
data$T3601400[361.0 <= data$T3601400 & data$T3601400 <= 380.0] <- 361.0
data$T3601400 <- factor(data$T3601400,
levels=c(0.0,201.0,221.0,241.0,261.0,281.0,301.0,321.0,341.0,361.0),
labels=c("0 TO 200: <200",
"201 TO 220",
"221 TO 240",
"241 TO 260",
"261 TO 280",
"281 TO 300",
"301 TO 320",
"321 TO 340",
"341 TO 360",
"361 TO 380"))
data$T3601500[30.0 <= data$T3601500 & data$T3601500 <= 31.0] <- 30.0
data$T3601500 <- factor(data$T3601500,
levels=c(24.0,25.0,26.0,27.0,28.0,29.0,30.0),
labels=c("24",
"25",
"26",
"27",
"28",
"29",
"30 TO 31: Age 30 or greater"))
data$T4565600[10.0 <= data$T4565600 & data$T4565600 <= 999.0] <- 10.0
data$T4565600 <- factor(data$T4565600,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 TO 999: 10+"))
data$T5201300[0.0 <= data$T5201300 & data$T5201300 <= 300.0] <- 0.0
data$T5201300[301.0 <= data$T5201300 & data$T5201300 <= 320.0] <- 301.0
data$T5201300[321.0 <= data$T5201300 & data$T5201300 <= 340.0] <- 321.0
data$T5201300[341.0 <= data$T5201300 & data$T5201300 <= 360.0] <- 341.0
data$T5201300[361.0 <= data$T5201300 & data$T5201300 <= 380.0] <- 361.0
data$T5201300 <- factor(data$T5201300,
levels=c(0.0,301.0,321.0,341.0,361.0),
labels=c("0 TO 300",
"301 TO 320",
"321 TO 340",
"341 TO 360",
"361 TO 380"))
data$T6209800[10.0 <= data$T6209800 & data$T6209800 <= 999.0] <- 10.0
data$T6209800 <- factor(data$T6209800,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 TO 999: 10+"))
data$T6651200[0.0 <= data$T6651200 & data$T6651200 <= 300.0] <- 0.0
data$T6651200[301.0 <= data$T6651200 & data$T6651200 <= 320.0] <- 301.0
data$T6651200[321.0 <= data$T6651200 & data$T6651200 <= 340.0] <- 321.0
data$T6651200[341.0 <= data$T6651200 & data$T6651200 <= 360.0] <- 341.0
data$T6651200[361.0 <= data$T6651200 & data$T6651200 <= 380.0] <- 361.0
data$T6651200[381.0 <= data$T6651200 & data$T6651200 <= 400.0] <- 381.0
data$T6651200 <- factor(data$T6651200,
levels=c(0.0,301.0,321.0,341.0,361.0,381.0),
labels=c("0 TO 300",
"301 TO 320",
"321 TO 340",
"341 TO 360",
"361 TO 380",
"381 TO 400"))
data$T7707200[10.0 <= data$T7707200 & data$T7707200 <= 999.0] <- 10.0
data$T7707200 <- factor(data$T7707200,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 TO 999: 10+"))
data$T8123500[0.0 <= data$T8123500 & data$T8123500 <= 340.0] <- 0.0
data$T8123500[341.0 <= data$T8123500 & data$T8123500 <= 360.0] <- 341.0
data$T8123500[361.0 <= data$T8123500 & data$T8123500 <= 380.0] <- 361.0
data$T8123500[381.0 <= data$T8123500 & data$T8123500 <= 400.0] <- 381.0
data$T8123500[401.0 <= data$T8123500 & data$T8123500 <= 420.0] <- 401.0
data$T8123500[421.0 <= data$T8123500 & data$T8123500 <= 440.0] <- 421.0
data$T8123500[441.0 <= data$T8123500 & data$T8123500 <= 460.0] <- 441.0
data$T8123500 <- factor(data$T8123500,
levels=c(0.0,341.0,361.0,381.0,401.0,421.0,441.0),
labels=c("0 TO 340",
"341 TO 360",
"361 TO 380",
"381 TO 400",
"401 TO 420",
"421 TO 440",
"441 TO 460"))
data$U0001700[0.0 <= data$U0001700 & data$U0001700 <= 340.0] <- 0.0
data$U0001700[341.0 <= data$U0001700 & data$U0001700 <= 360.0] <- 341.0
data$U0001700[361.0 <= data$U0001700 & data$U0001700 <= 380.0] <- 361.0
data$U0001700[381.0 <= data$U0001700 & data$U0001700 <= 400.0] <- 381.0
data$U0001700[401.0 <= data$U0001700 & data$U0001700 <= 420.0] <- 401.0
data$U0001700[421.0 <= data$U0001700 & data$U0001700 <= 440.0] <- 421.0
data$U0001700[441.0 <= data$U0001700 & data$U0001700 <= 460.0] <- 441.0
data$U0001700 <- factor(data$U0001700,
levels=c(0.0,341.0,361.0,381.0,401.0,421.0,441.0),
labels=c("0 TO 340",
"341 TO 360",
"361 TO 380",
"381 TO 400",
"401 TO 420",
"421 TO 440",
"441 TO 460"))
data$U1099900[10.0 <= data$U1099900 & data$U1099900 <= 999.0] <- 10.0
data$U1099900 <- factor(data$U1099900,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 TO 999: 10+"))
data$U1838400[0.0 <= data$U1838400 & data$U1838400 <= 340.0] <- 0.0
data$U1838400[341.0 <= data$U1838400 & data$U1838400 <= 360.0] <- 341.0
data$U1838400[361.0 <= data$U1838400 & data$U1838400 <= 380.0] <- 361.0
data$U1838400[381.0 <= data$U1838400 & data$U1838400 <= 400.0] <- 381.0
data$U1838400[401.0 <= data$U1838400 & data$U1838400 <= 420.0] <- 401.0
data$U1838400[421.0 <= data$U1838400 & data$U1838400 <= 440.0] <- 421.0
data$U1838400[441.0 <= data$U1838400 & data$U1838400 <= 460.0] <- 441.0
data$U1838400[461.0 <= data$U1838400 & data$U1838400 <= 480.0] <- 461.0
data$U1838400 <- factor(data$U1838400,
levels=c(0.0,341.0,361.0,381.0,401.0,421.0,441.0,461.0),
labels=c("0 TO 340",
"341 TO 360",
"361 TO 380",
"381 TO 400",
"401 TO 420",
"421 TO 440",
"441 TO 460",
"461 TO 480"))
data$U2963000[10.0 <= data$U2963000 & data$U2963000 <= 999.0] <- 10.0
data$U2963000 <- factor(data$U2963000,
levels=c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0),
labels=c("0",
"1",
"2",
"3",
"4",
"5",
"6",
"7",
"8",
"9",
"10 TO 999: 10+"))
data$U3437800[0.0 <= data$U3437800 & data$U3437800 <= 340.0] <- 0.0
data$U3437800[341.0 <= data$U3437800 & data$U3437800 <= 360.0] <- 341.0
data$U3437800[361.0 <= data$U3437800 & data$U3437800 <= 380.0] <- 361.0
data$U3437800[381.0 <= data$U3437800 & data$U3437800 <= 400.0] <- 381.0
data$U3437800[401.0 <= data$U3437800 & data$U3437800 <= 420.0] <- 401.0
data$U3437800[421.0 <= data$U3437800 & data$U3437800 <= 440.0] <- 421.0
data$U3437800[441.0 <= data$U3437800 & data$U3437800 <= 460.0] <- 441.0
data$U3437800[461.0 <= data$U3437800 & data$U3437800 <= 480.0] <- 461.0
data$U3437800[481.0 <= data$U3437800 & data$U3437800 <= 500.0] <- 481.0
data$U3437800[501.0 <= data$U3437800 & data$U3437800 <= 520.0] <- 501.0
data$U3437800 <- factor(data$U3437800,
levels=c(0.0,341.0,361.0,381.0,401.0,421.0,441.0,461.0,481.0,501.0),
labels=c("0 TO 340",
"341 TO 360",
"361 TO 380",
"381 TO 400",
"401 TO 420",
"421 TO 440",
"441 TO 460",
"461 TO 480",
"481 TO 500",
"501 TO 520"))
data$U4942900[0.0 <= data$U4942900 & data$U4942900 <= 340.0] <- 0.0
data$U4942900[341.0 <= data$U4942900 & data$U4942900 <= 360.0] <- 341.0
data$U4942900[361.0 <= data$U4942900 & data$U4942900 <= 380.0] <- 361.0
data$U4942900[381.0 <= data$U4942900 & data$U4942900 <= 400.0] <- 381.0
data$U4942900[401.0 <= data$U4942900 & data$U4942900 <= 420.0] <- 401.0
data$U4942900[421.0 <= data$U4942900 & data$U4942900 <= 440.0] <- 421.0
data$U4942900[441.0 <= data$U4942900 & data$U4942900 <= 460.0] <- 441.0
data$U4942900[461.0 <= data$U4942900 & data$U4942900 <= 480.0] <- 461.0
data$U4942900[481.0 <= data$U4942900 & data$U4942900 <= 500.0] <- 481.0
data$U4942900[501.0 <= data$U4942900 & data$U4942900 <= 520.0] <- 501.0
data$U4942900 <- factor(data$U4942900,
levels=c(0.0,341.0,361.0,381.0,401.0,421.0,441.0,461.0,481.0,501.0),
labels=c("0 TO 340",
"341 TO 360",
"361 TO 380",
"381 TO 400",
"401 TO 420",
"421 TO 440",
"441 TO 460",
"461 TO 480",
"481 TO 500",
"501 TO 520"))
data$U5868001[1.0 <= data$U5868001 & data$U5868001 <= 2.0] <- 1.0
data$U5868001 <- factor(data$U5868001,
levels=c(0.0,1.0),
labels=c("NOT SELECTED",
"Broadband (high speed) internet service such cable, fiber optic, or DSL service installed in your home"))
data$U5868002[1.0 <= data$U5868002 & data$U5868002 <= 3.0] <- 1.0
data$U5868002 <- factor(data$U5868002,
levels=c(0.0,1.0),
labels=c("NOT SELECTED",
"Satellite internet service installed in your home"))
data$U5868003[1.0 <= data$U5868003 & data$U5868003 <= 4.0] <- 1.0
data$U5868003 <- factor(data$U5868003,
levels=c(0.0,1.0),
labels=c("NOT SELECTED",
"Dial-up internet service installed in your home"))
data$U5868004[1.0 <= data$U5868004 & data$U5868004 <= 5.0] <- 1.0
data$U5868004 <- factor(data$U5868004,
levels=c(0.0,1.0),
labels=c("NOT SELECTED",
"Some other service"))
return(data)
}

varlabels <- c("PUBID - YTH ID CODE 1997",
"HOW RS GENL HEALTH? 1997",
"# DAYS/SCH WK R EATS BREAKFAST 1997",
"# DAYS/WK EAT VEGGIES/FRUIT? 1997",
"# DAYS/WK EXERCISE 30+ MINUTES 1997",
"% TIME WEAR SEATBELT IN CAR? 1997",
"ENT % TIME WEAR SEATBELT 1997",
"KEY!SEX (SYMBOL) 1997",
"KEY!BDATE M/Y (SYMBOL) 1997",
"KEY!BDATE M/Y (SYMBOL) 1997",
"KEY!AGE (SYMBOL) 1997",
"CV_AGE(MONTHS)_INT_DATE 1997",
"CV_AGE_12/31/96 1997",
"CV_AGE_INT_DATE 1997",
"CV_BIO_MOM_AGE_CHILD1 1997",
"CV_BIO_MOM_AGE_YOUTH 1997",
"CV_CENSUS_REGION 1997",
"CV_CITIZENSHIP 1997",
"CV_MSA 1997",
"CV_URBAN-RURAL 1997",
"CV_SAMPLE_TYPE 1997",
"CV_HGC_BIO_DAD 1997",
"CV_HGC_BIO_MOM 1997",
"CV_HGC_RES_DAD 1997",
"CV_HGC_RES_MOM 1997",
"KEY!RACE_ETHNICITY (SYMBOL) 1997",
"HOW RS GENERAL HEALTH? 1998",
"R OWN, RENT HOUSE/APT? 1998",
"WHAT HOUSE/APT OWNER SITUATION? 1998",
"CV_AGE(MONTHS)_INT_DATE 1998",
"CV_AGE_INT_DATE 1998",
"CV_CENSUS_REGION 1998",
"CV_MSA 1998",
"CV_URBAN-RURAL 1998",
"HOW RS GENERAL HEALTH? 1999",
"R OWN, RENT HOUSE/APT? 1999",
"WHAT HOUSE/APT OWNER SITUATION? 1999",
"CV_AGE(MONTHS)_INT_DATE 1999",
"CV_AGE_INT_DATE 1999",
"CV_CENSUS_REGION 1999",
"CV_MSA 1999",
"CV_URBAN-RURAL 1999",
"HOW RS GENERAL HEALTH? 2000",
"HOW OFT R BEEN NERVOUS PERSON 2000",
"HOW OFT R CALM/PEACEFUL PAST MO 2000",
"HOW OFT R FELT DOWN OR BLUE 2000",
"HOW OFT R BEEN HAPPY PERSON 2000",
"HOW OFT R DEPRESSED LAST MONTH 2000",
"R OWN, RENT HOUSE/APT? 2000",
"WHAT HOUSE/APT OWNER SITUATION? 2000",
"CV_AGE(MONTHS)_INT_DATE 2000",
"CV_AGE_INT_DATE 2000",
"CV_CENSUS_REGION 2000",
"CV_MSA 2000",
"CV_URBAN-RURAL 2000",
"HOW RS GENERAL HEALTH? 2001",
"R OWN, RENT HOUSE/APT? 2001",
"WHAT HOUSE/APT OWNER SITUATION? 2001",
"CV_AGE(MONTHS)_INT_DATE 2001",
"CV_AGE_INT_DATE 2001",
"CV_CENSUS_REGION 2001",
"CV_MSA 2001",
"CV_URBAN-RURAL 2001",
"CV_CITIZENSHIP_CURR 2001",
"HOW OFT R BEEN NERVOUS PERSON 2002",
"HOW OFT R CALM/PEACEFUL PAST MO 2002",
"HOW OFT R FELT DOWN OR BLUE 2002",
"HOW OFT R BEEN HAPPY PERSON 2002",
"HOW OFT R DEPRESSED LAST MONTH 2002",
"R OWN, RENT HOUSE/APT? 2002",
"WHAT HOUSE/APT OWNER SITUATION? 2002",
"HOW R'S GENERAL HEALTH? 2002",
"TIMES/WK R EAT FRUIT 2002",
"TIMES/WK R EAT VEGETABLES 2002",
"TIMES/WK R EXERCISE 30+ MIN 2002",
"HRS/WK R USES A COMPUTER 2002",
"HRS/WK R WATCHES TELEVISION 2002",
"HRS/NIGHT R SLEEPS 2002",
"TIMES R DRIVE WHEN DRINKING 2002",
"TIMES RIDE W/SOMEONE DRINKING 2002",
"CV_AGE(MONTHS)_INT_DATE 2002",
"CV_AGE_INT_DATE 2002",
"CV_CENSUS_REGION 2002",
"CV_CITIZENSHIP_CURR 2002",
"CV_MSA 2002",
"CV_URBAN-RURAL 2002",
"CV_AGE(MONTHS)_INT_DATE 2003",
"CV_AGE_INT_DATE 2003",
"CV_CENSUS_REGION 2003",
"CV_CITIZENSHIP_CURR 2003",
"CV_MSA 2003",
"CV_URBAN-RURAL 2003",
"R OWN, RENT HOUSE/APT? 2003",
"WHAT HOUSE/APT OWNER SITUATION? 2003",
"HOW R'S GENERAL HEALTH? 2003",
"CV_AGE(MONTHS)_INT_DATE 2004",
"CV_AGE_INT_DATE 2004",
"CV_CENSUS_REGION 2004",
"CV_CITIZENSHIP_CURR 2004",
"CV_MSA 2004",
"CV_URBAN-RURAL 2004",
"HOW OFT R BEEN NERVOUS PERSON 2004",
"HOW OFT R CALM/PEACEFUL PAST MO 2004",
"HOW OFT R FELT DOWN OR BLUE 2004",
"HOW OFT R BEEN HAPPY PERSON 2004",
"HOW OFT R DEPRESSED LAST MONTH 2004",
"R OWN, RENT HOUSE/APT? 2004",
"WHAT HOUSE/APT OWNER SITUATION? 2004",
"HOW R'S GENERAL HEALTH? 2004",
"CV_AGE(MONTHS)_INT_DATE 2005",
"CV_AGE_INT_DATE 2005",
"CV_CENSUS_REGION 2005",
"CV_MSA 2005",
"CV_URBAN-RURAL 2005",
"R OWN, RENT HOUSE/APT? 2005",
"WHAT HOUSE/APT OWNER SITUATION? 2005",
"HOW R'S GENERAL HEALTH? 2005",
"CV_AGE(MONTHS)_INT_DATE 2006",
"CV_AGE_INT_DATE 2006",
"CV_CENSUS_REGION 2006",
"CV_CITIZENSHIP_CURR 2006",
"CV_MSA 2006",
"CV_URBAN-RURAL 2006",
"HOW OFT R BEEN NERVOUS PERSON 2006",
"HOW OFT R CALM/PEACEFUL PAST MO 2006",
"HOW OFT R FELT DOWN OR BLUE 2006",
"HOW OFT R BEEN HAPPY PERSON 2006",
"HOW OFT R DEPRESSED LAST MONTH 2006",
"R OWN, RENT HOUSE/APT? 2006",
"WHAT HOUSE/APT OWNER SITUATION? 2006",
"HOW R'S GENERAL HEALTH? 2006",
"HOW EXCITING IS RS LIFE? 2006",
"HOW R RATES HIS CURRENT LIFE 2006",
"CV_AGE(MONTHS)_INT_DATE 2007",
"CV_AGE_INT_DATE 2007",
"CV_CENSUS_REGION 2007",
"CV_CITIZENSHIP_CURR 2007",
"CV_MSA 2007",
"CV_URBAN-RURAL 2007",
"R OWN, RENT HOUSE/APT? 2007",
"R OWN, RENT HOUSE/APT? 2007",
"WHAT HOUSE/APT OWNER SITUATION? 2007",
"HOW R'S GENERAL HEALTH? 2007",
"TIMES/WK R EAT FRUIT 2007",
"TIMES/WK R EAT VEGETABLES 2007",
"TIMES/WK R EXERCISE 30+ MIN 2007",
"HRS/WK R USES A COMPUTER 2007",
"HRS/WK R WATCHES TELEVISION 2007",
"HRS/NIGHT R SLEEPS 2007",
"TIMES R DRIVE WHEN DRINKING 2007",
"TIMES RIDE W/SOMEONE DRINKING 2007",
"HOW R RATES HIS CURRENT LIFE 2007",
"CV_AGE(MONTHS)_INT_DATE 2008",
"CV_AGE_INT_DATE 2008",
"CV_CENSUS_REGION 2008",
"CV_CITIZENSHIP_CURR 2008",
"CV_MSA 2008",
"CV_URBAN-RURAL 2008",
"R OWN, RENT HOUSE/APT? 2008",
"HOW OFT R BEEN NERVOUS PERSON 2008",
"HOW OFT R CALM/PEACEFUL PAST MO 2008",
"HOW OFT R FELT DOWN OR BLUE 2008",
"HOW OFT R BEEN HAPPY PERSON 2008",
"HOW OFT R DEPRESSED LAST MONTH 2008",
"R OWN, RENT HOUSE/APT? 2008",
"WHAT HOUSE/APT OWNER SITUATION? 2008",
"HOW R'S GENERAL HEALTH? 2008",
"TIMES/WK R EAT FRUIT 2008",
"TIMES/WK R EAT VEGETABLES 2008",
"TIMES/WK R EXERCISE 30+ MIN 2008",
"HRS/WK R USES A COMPUTER 2008",
"HRS/WK R WATCHES TELEVISION 2008",
"HRS/NIGHT R SLEEPS 2008",
"TIMES R DRIVE WHEN DRINKING 2008",
"TIMES RIDE W/SOMEONE DRINKING 2008",
"HOW R RATES HIS CURRENT LIFE 2008",
"CV_AGE(MONTHS)_INT_DATE 2009",
"CV_AGE_INT_DATE 2009",
"CV_CENSUS_REGION 2009",
"CV_CITIZENSHIP_CURR 2009",
"CV_MSA 2009",
"CV_URBAN-RURAL 2009",
"R OWN, RENT HOUSE/APT? 2009",
"HOW R'S GENERAL HEALTH? 2009",
"TIMES/WK R EAT FRUIT 2009",
"TIMES/WK R EAT VEGETABLES 2009",
"HRS/WK R USES A COMPUTER 2009",
"HRS/WK R WATCHES TELEVISION 2009",
"HRS/NIGHT R SLEEPS 2009",
"TIMES R DRIVE WHEN DRINKING 2009",
"TIMES RIDE W/SOMEONE DRINKING 2009",
"CV_AGE(MONTHS)_INT_DATE 2010",
"CV_AGE_INT_DATE 2010",
"CV_CENSUS_REGION 2010",
"CV_CITIZENSHIP_CURR 2010",
"CV_MSA 2010",
"CV_URBAN-RURAL 2010",
"HOW OFT R BEEN NERVOUS PERSON 2010",
"HOW OFT R CALM/PEACEFUL PAST MO 2010",
"HOW OFT R FELT DOWN OR BLUE 2010",
"HOW OFT R BEEN HAPPY PERSON 2010",
"HOW OFT R DEPRESSED LAST MONTH 2010",
"HOW R'S GENERAL HEALTH? 2010",
"TIMES/WK R EAT FRUIT 2010",
"TIMES/WK R EAT VEGETABLES 2010",
"HRS/WK R USES A COMPUTER 2010",
"HRS/WK R WATCHES TELEVISION 2010",
"HRS/NIGHT R SLEEPS 2010",
"TIMES R DRIVE WHEN DRINKING 2010",
"TIMES RIDE W/SOMEONE DRINKING 2010",
"CV_AGE(MONTHS)_INT_DATE 2011",
"CV_AGE_INT_DATE 2011",
"CV_CENSUS_REGION 2011",
"CV_CITIZENSHIP_CURR 2011",
"CV_MSA 2011",
"CV_URBAN-RURAL 2011",
"HOW R'S GENERAL HEALTH? 2011",
"TIMES/WK R EAT FRUIT 2011",
"TIMES/WK R EAT VEGETABLES 2011",
"HRS/WK R USES A COMPUTER 2011",
"HRS/WK R WATCHES TELEVISION 2011",
"HRS/NIGHT R SLEEPS 2011",
"TIMES R DRIVE WHEN DRINKING 2011",
"TIMES RIDE W/SOMEONE DRINKING 2011",
"CV_AGE(MONTHS)_INT_DATE 2013",
"CV_AGE_INT_DATE 2013",
"CV_CENSUS_REGION 2013",
"CV_CITIZENSHIP_CURR 2013",
"CV_MSA 2013",
"CV_URBAN-RURAL 2013",
"HOW R'S GENERAL HEALTH? 2013",
"CV_AGE(MONTHS)_INT_DATE 2015",
"CV_AGE_INT_DATE 2015",
"CV_CENSUS_REGION 2015",
"CV_CITIZENSHIP_CURR 2015",
"CV_MSA 2015",
"CV_URBAN-RURAL 2015",
"HOW OFT R BEEN NERVOUS PERSON 2015",
"HOW OFT R CALM/PEACEFUL PAST MO 2015",
"HOW OFT R FELT DOWN OR BLUE 2015",
"HOW OFT R BEEN HAPPY PERSON 2015",
"HOW OFT R DEPRESSED LAST MONTH 2015",
"HOW R'S GENERAL HEALTH? 2015",
"TIMES/WK R EAT FRUIT 2015",
"TIMES/WK R EAT VEGETABLES 2015",
"HRS/WK R USES A COMPUTER 2015",
"HRS/WK R WATCHES TELEVISION 2015",
"HRS/NIGHT R SLEEPS 2015",
"TIMES R DRIVE WHEN DRINKING 2015",
"TIMES RIDE W/SOMEONE DRINKING 2015",
"CV_AGE(MONTHS)_INT_DATE 2017",
"CV_AGE_INT_DATE 2017",
"CV_CENSUS_REGION 2017",
"CV_MSA 2017",
"CV_URBAN-RURAL 2017",
"HOW R'S GENERAL HEALTH? 2017",
"HRS/WK R USES A COMPUTER 2017",
"HRS/WK R WATCHES TELEVISION 2017",
"HRS/NIGHT R SLEEPS 2017",
"TIMES R DRIVE WHEN DRINKING 2017",
"TIMES RIDE W/SOMEONE DRINKING 2017",
"HOW OFT R BEEN NERVOUS PERSON 2017",
"HOW OFT R CALM/PEACEFUL PAST MO 2017",
"HOW OFT R FELT DOWN OR BLUE 2017",
"HOW OFT R BEEN HAPPY PERSON 2017",
"HOW OFT R DEPRESSED LAST MONTH 2017",
"CV_AGE(MONTHS)_INT_DATE 2019",
"CV_AGE_INT_DATE 2019",
"CV_CENSUS_REGION 2019",
"CV_MSA 2019",
"CV_URBAN-RURAL 2019",
"HOW R'S GENERAL HEALTH? 2019",
"HOW OFT R BEEN NERVOUS PERSON 2019",
"HOW OFT R CALM/PEACEFUL PAST MO 2019",
"HOW OFT R FELT DOWN OR BLUE 2019",
"HOW OFT R BEEN HAPPY PERSON 2019",
"HOW OFT R DEPRESSED LAST MONTH 2019",
"CV_AGE(MONTHS)_INT_DATE 2021",
"CV_AGE_INT_DATE 2021",
"CV_CENSUS_REGION 2021",
"CV_MSA 2021",
"CV_URBAN-RURAL 2021",
"HOW R'S GENERAL HEALTH? 2021",
"INTERNET ACCESS AT HOME? 2021",
"TYPE OF HOME INTERNET ACCESS - CELL PLAN 2021",
"TYPE OF HOME INTERNET ACCESS - BROADBAND 2021",
"TYPE OF HOME INTERNET ACCESS - SATELLITE 2021",
"TYPE OF HOME INTERNET ACCESS - DIALUP 2021",
"TYPE OF HOME INTERNET ACCESS - OTHER 2021",
"HOME INTERNET ACCESS - TYPE OF DEVICE 2021",
"R OWN, RENT HOUSE/APT?",
"WHAT HOUSE/APT OWNER SITUATION?",
"HEA29: RS ENERGY LEVEL PAST 4 WKS"
)


# Use qnames rather than rnums

qnames = function(data) {
names(data) <- c("PUBID_1997",
"YHEA-100_1997",
"YHEA-400_1997",
"YHEA-500_1997",
"YHEA-600_1997",
"YHEA-700_1997",
"YHEA-800_1997",
"KEY_SEX_1997",
"KEY_BDATE_M_1997",
"KEY_BDATE_Y_1997",
"KEY_AGE_1997",
"CV_AGE(MONTHS)_INT_DATE_1997",
"CV_AGE_12/31/96_1997",
"CV_AGE_INT_DATE_1997",
"CV_BIO_MOM_AGE_CHILD1_1997",
"CV_BIO_MOM_AGE_YOUTH_1997",
"CV_CENSUS_REGION_1997",
"CV_CITIZENSHIP_1997",
"CV_MSA_1997",
"CV_URBAN-RURAL_1997",
"CV_SAMPLE_TYPE_1997",
"CV_HGC_BIO_DAD_1997",
"CV_HGC_BIO_MOM_1997",
"CV_HGC_RES_DAD_1997",
"CV_HGC_RES_MOM_1997",
"KEY_RACE_ETHNICITY_1997",
"YHEA-100_1998",
"YAST-3310_1998",
"YAST-3320_1998",
"CV_AGE(MONTHS)_INT_DATE_1998",
"CV_AGE_INT_DATE_1998",
"CV_CENSUS_REGION_1998",
"CV_MSA_1998",
"CV_URBAN-RURAL_1998",
"YHEA-100_1999",
"YAST-3310_1999",
"YAST-3320_1999",
"CV_AGE(MONTHS)_INT_DATE_1999",
"CV_AGE_INT_DATE_1999",
"CV_CENSUS_REGION_1999",
"CV_MSA_1999",
"CV_URBAN-RURAL_1999",
"YHEA-100_2000",
"YSAQ-282C_2000",
"YSAQ-282D_2000",
"YSAQ-282E_2000",
"YSAQ-282F_2000",
"YSAQ-282G_2000",
"YAST-3310_2000",
"YAST-3320_2000",
"CV_AGE(MONTHS)_INT_DATE_2000",
"CV_AGE_INT_DATE_2000",
"CV_CENSUS_REGION_2000",
"CV_MSA_2000",
"CV_URBAN-RURAL_2000",
"YHEA-100_2001",
"YAST-3310_2001",
"YAST-3320_2001",
"CV_AGE(MONTHS)_INT_DATE_2001",
"CV_AGE_INT_DATE_2001",
"CV_CENSUS_REGION_2001",
"CV_MSA_2001",
"CV_URBAN-RURAL_2001",
"CV_CITIZEN_CURRENT_2001",
"YSAQ-282C_2002",
"YSAQ-282D_2002",
"YSAQ-282E_2002",
"YSAQ-282F_2002",
"YSAQ-282G_2002",
"YAST-3310_2002",
"YAST-3320_2002",
"YHEA-100_2002",
"YHEA-500A_2002",
"YHEA-500B_2002",
"YHEA-600_2002",
"YHEA-610_2002",
"YHEA-620_2002",
"YHEA-650_2002",
"YHEA-810_2002",
"YHEA-820_2002",
"CV_AGE(MONTHS)_INT_DATE_2002",
"CV_AGE_INT_DATE_2002",
"CV_CENSUS_REGION_2002",
"CV_CITIZEN_CURRENT_2002",
"CV_MSA_2002",
"CV_URBAN-RURAL_2002",
"CV_AGE(MONTHS)_INT_DATE_2003",
"CV_AGE_INT_DATE_2003",
"CV_CENSUS_REGION_2003",
"CV_CITIZEN_CURRENT_2003",
"CV_MSA_2003",
"CV_URBAN-RURAL_2003",
"YAST-3310_2003",
"YAST-3320_2003",
"YHEA-100_2003",
"CV_AGE_MONTHS_INT_DATE_2004",
"CV_AGE_INT_DATE_2004",
"CV_CENSUS_REGION_2004",
"CV_CITIZEN_CURRENT_2004",
"CV_MSA_2004",
"CV_URBAN-RURAL_2004",
"YSAQ-282C_2004",
"YSAQ-282D_2004",
"YSAQ-282E_2004",
"YSAQ-282F_2004",
"YSAQ-282G_2004",
"YAST-3310_2004",
"YAST-3320_2004",
"YHEA-100_2004",
"CV_AGE_MONTHS_INT_DATE_2005",
"CV_AGE_INT_DATE_2005",
"CV_CENSUS_REGION_2005",
"CV_MSA_2005",
"CV_URBAN-RURAL_2005",
"YAST-3310_2005",
"YAST-3320_2005",
"YHEA-100_2005",
"CV_AGE_MONTHS_INT_DATE_2006",
"CV_AGE_INT_DATE_2006",
"CV_CENSUS_REGION_2006",
"CV_CITIZEN_CURRENT_2006",
"CV_MSA_2006",
"CV_URBAN-RURAL_2006",
"YSAQ-282C_2006",
"YSAQ-282D_2006",
"YSAQ-282E_2006",
"YSAQ-282F_2006",
"YSAQ-282G_2006",
"YAST-3310_2006",
"YAST-3320_2006",
"YHEA-100_2006",
"YTEL-4_2006",
"YTEL-21_2006",
"CV_AGE_MONTHS_INT_DATE_2007",
"CV_AGE_INT_DATE_2007",
"CV_CENSUS_REGION_2007",
"CV_CITIZEN_CURRENT_2007",
"CV_MSA_2007",
"CV_URBAN-RURAL_2007",
"YHHI-AST-3310_2007",
"YAST-3310_2007",
"YAST-3320_2007",
"YHEA-100_2007",
"YHEA-500A_2007",
"YHEA-500B_2007",
"YHEA-600_2007",
"YHEA-610_2007",
"YHEA-620_2007",
"YHEA-650_2007",
"YHEA-810_2007",
"YHEA-820_2007",
"YTEL-21_2007",
"CV_AGE_MONTHS_INT_DATE_2008",
"CV_AGE_INT_DATE_2008",
"CV_CENSUS_REGION_2008",
"CV_CITIZEN_CURRENT_2008",
"CV_MSA_2008",
"CV_URBAN-RURAL_2008",
"YHHI-AST-3310_2008",
"YSAQ-282C_2008",
"YSAQ-282D_2008",
"YSAQ-282E_2008",
"YSAQ-282F_2008",
"YSAQ-282G_2008",
"YAST-3310_2008",
"YAST-3320_2008",
"YHEA-100_2008",
"YHEA-500A_2008",
"YHEA-500B_2008",
"YHEA-600_2008",
"YHEA-610_2008",
"YHEA-620_2008",
"YHEA-650_2008",
"YHEA-810_2008",
"YHEA-820_2008",
"YTEL-55_2008",
"CV_AGE_MONTHS_INT_DATE_2009",
"CV_AGE_INT_DATE_2009",
"CV_CENSUS_REGION_2009",
"CV_CITIZEN_CURRENT_2009",
"CV_MSA_2009",
"CV_URBAN-RURAL_2009",
"YHHI-AST-3310_2009",
"YHEA-100_2009",
"YHEA-500A_2009",
"YHEA-500B_2009",
"YHEA-610_2009",
"YHEA-620_2009",
"YHEA-650_2009",
"YHEA-810_2009",
"YHEA-820_2009",
"CV_AGE_MONTHS_INT_DATE_2010",
"CV_AGE_INT_DATE_2010",
"CV_CENSUS_REGION_2010",
"CV_CITIZEN_CURRENT_2010",
"CV_MSA_2010",
"CV_URBAN-RURAL_2010",
"YSAQ-282C_2010",
"YSAQ-282D_2010",
"YSAQ-282E_2010",
"YSAQ-282F_2010",
"YSAQ-282G_2010",
"YHEA-100_2010",
"YHEA-500A_2010",
"YHEA-500B_2010",
"YHEA-610_2010",
"YHEA-620_2010",
"YHEA-650_2010",
"YHEA-810_2010",
"YHEA-820_2010",
"CV_AGE_MONTHS_INT_DATE_2011",
"CV_AGE_INT_DATE_2011",
"CV_CENSUS_REGION_2011",
"CV_CITIZEN_CURRENT_2011",
"CV_MSA_2011",
"CV_URBAN-RURAL_2011",
"YHEA-100_2011",
"YHEA-500A_2011",
"YHEA-500B_2011",
"YHEA-610_2011",
"YHEA-620_2011",
"YHEA-650_2011",
"YHEA-810_2011",
"YHEA-820_2011",
"CV_AGE_MONTHS_INT_DATE_2013",
"CV_AGE_INT_DATE_2013",
"CV_CENSUS_REGION_2013",
"CV_CITIZEN_CURRENT_2013",
"CV_MSA_2013",
"CV_URBAN-RURAL_2013",
"YHEA-100_2013",
"CV_AGE_MONTHS_INT_DATE_2015",
"CV_AGE_INT_DATE_2015",
"CV_CENSUS_REGION_2015",
"CV_CITIZEN_CURRENT_2015",
"CV_MSA_2015",
"CV_URBAN-RURAL_2015",
"YSAQ-282C_2015",
"YSAQ-282D_2015",
"YSAQ-282E_2015",
"YSAQ-282F_2015",
"YSAQ-282G_2015",
"YHEA-100_2015",
"YHEA-500A_2015",
"YHEA-500B_2015",
"YHEA-610_2015",
"YHEA-620_2015",
"YHEA-650_2015",
"YHEA-810_2015",
"YHEA-820_2015",
"CV_AGE_MONTHS_INT_DATE_2017",
"CV_AGE_INT_DATE_2017",
"CV_CENSUS_REGION_2017",
"CV_MSA_2017",
"CV_URBAN-RURAL_2017",
"YHEA-100_2017",
"YHEA-610_2017",
"YHEA-620_2017",
"YHEA-650_2017",
"YHEA-810_2017",
"YHEA-820_2017",
"YHEA-SAQ-282C_2017",
"YHEA-SAQ-282D_2017",
"YHEA-SAQ-282E_2017",
"YHEA-SAQ-282F_2017",
"YHEA-SAQ-282G_2017",
"CV_AGE_MONTHS_INT_DATE_2019",
"CV_AGE_INT_DATE_2019",
"CV_CENSUS_REGION_2019",
"CV_MSA_2019",
"CV_URBAN-RURAL_2019",
"YHEA-100_2019",
"YHEA-SAQ-282C_2019",
"YHEA-SAQ-282D_2019",
"YHEA-SAQ-282E_2019",
"YHEA-SAQ-282F_2019",
"YHEA-SAQ-282G_2019",
"CV_AGE_MONTHS_INT_DATE_2021",
"CV_AGE_INT_DATE_2021",
"CV_CENSUS_REGION_2021",
"CV_MSA_2021",
"CV_URBAN-RURAL_2021",
"YHEA-100_2021",
"YEND-INTERNET1_2021",
"YEND-INTERNET2~000001_2021",
"YEND-INTERNET2~000002_2021",
"YEND-INTERNET2~000003_2021",
"YEND-INTERNET2~000004_2021",
"YEND-INTERNET2~000005_2021",
"YEND-INTERNET4_2021",
"YAST25-3310_COMB_XRND",
"YAST25-3320_COMB_XRND",
"YHEA29-285_COMB_XRND")
return(data)
}


#********************************************************************************************************

# Remove the '#' before the following line to create a data file called "categories" with value labels.
categories <- vallabels(new_data)

# Remove the '#' before the following lines to rename variables using Qnames instead of Reference Numbers
new_data <- qnames(new_data)
#categories <- qnames(categories)

# Produce summaries for the raw (uncategorized) data file
summary(new_data)

# Remove the '#' before the following lines to produce summaries for the "categories" data file.
#categories <- vallabels(new_data)
#categories <- vallabels_continuous(new_data)
#summary(categories)

#************************************************************************************************************

```

```{r wrangle_and_select, eval = FALSE, include = FALSE}
data0 <- new_data

data_named <- data0
colnames(data_named) <- varlabels


data_named[1:10, c(9, 10, 140, 141,159, 165) ]
#colnames(data_named) # keep 10, 140, 159, get rid of 9, 141, 165
data_named <- data_named[,-c(9, 141, 165)]
data_named <- data_named %>% 
  janitor::clean_names()

# Step 1: Pivot to a long format, separating the variable name from the year
df_long <- data_named %>%
  pivot_longer(
    cols = -pubid_yth_id_code_1997,  # Keep subject_id as a unique identifier
    names_to = c("variable", "year"),
    names_pattern = "^(.*)_(\\d{4})$",
    values_to = "value"
  ) %>% 
  mutate(id = pubid_yth_id_code_1997) %>% 
  select(-pubid_yth_id_code_1997) 


df_wide <- df_long %>%
  # Step 1: Remove rows with NA in critical columns
  filter(!is.na(value) & !is.na(year) & !is.na(variable) & !is.na(id)) %>%
  pivot_wider(
    names_from = variable,
    values_from = value
  ) 

# Select variables of interest and fix, recode health and emotions variables
df_select <- df_wide %>% 
  select(id, year, how_rs_genl_health, how_rs_general_health, cv_age_int_date, 
         key_sex_symbol, key_bdate_m_y_symbol, key_race_ethnicity_symbol,
         #  sex, race, cohort, 
         cv_urban_rural,
         hrs_wk_r_uses_a_computer, hrs_wk_r_watches_television, number_days_wk_exercise_30_minutes,
         how_r_rates_his_current_life, how_exciting_is_rs_life,
         how_oft_r_been_happy_person, how_oft_r_been_nervous_person, how_oft_r_calm_peaceful_past_mo, how_oft_r_felt_down_or_blue, how_oft_r_depressed_last_month) %>% 
  mutate(health = if_else(year == 1997, how_rs_genl_health, how_rs_general_health)) %>% 
  mutate(health = 6 - health) %>% 
  select(-c("how_rs_genl_health", "how_rs_general_health")) %>% 
  mutate(age = cv_age_int_date) %>% 
  mutate(felt_happy = 5 - how_oft_r_been_happy_person) %>% 
  mutate(felt_nervous = 5 - how_oft_r_been_nervous_person) %>% 
  mutate(felt_calm = 5 - how_oft_r_calm_peaceful_past_mo) %>% 
  mutate(felt_down = 5 - how_oft_r_felt_down_or_blue) %>% 
  mutate(depressed = 5 - how_oft_r_depressed_last_month) %>% 
  mutate(watched_tv = hrs_wk_r_watches_television) %>% 
  mutate(used_computer = hrs_wk_r_uses_a_computer) %>% 
  mutate(exercised = number_days_wk_exercise_30_minutes) %>% 
  mutate(rate_life = how_r_rates_his_current_life) %>% 
  select(c(id, year, age, key_sex_symbol, key_bdate_m_y_symbol, key_race_ethnicity_symbol,
           health, felt_happy, felt_nervous, felt_calm, felt_down, depressed, watched_tv, used_computer, exercised, rate_life))

# Put static variables from 1997 in every year
df_select <- df_select %>% 
  group_by(id) %>%
  mutate(sex = first(key_sex_symbol[year == 1997])) %>%
  mutate(cohort = first(key_bdate_m_y_symbol[year == 1997])) %>%
  mutate(race = first(key_race_ethnicity_symbol[year == 1997])) %>%
  ungroup() %>% 
  select(- c(key_sex_symbol, key_bdate_m_y_symbol, key_race_ethnicity_symbol)) %>% 
  select(id, year, health, sex, cohort, race, everything())

# Select only variables of interest and waves of interest
df_select <- df_select %>% 
  select(c(id, year, health, sex, cohort, race, age, health, 
           felt_happy, felt_nervous, felt_calm, felt_down, depressed, 
           watched_tv, used_computer
           )) %>% 
  na.omit()

write_csv(df_select, "data/panel/nlsy97_selected.csv")

# Check attrition
df_select %>%
  group_by(year) %>%
  drop_na(health) %>% 
  summarize(count = n_distinct(id))

last_wave <- df_select %>%
  group_by(id) %>%
  summarize(last_wave_present = max(year))

dropouts <- last_wave %>%
  filter(last_wave_present < max(last_wave_present))

```



# Load and explore data

```{r}

data <- read_csv("data/panel/nlsy97_selected.csv")  %>% 
  filter(year != 2008) # there were not enough observations in this wave

# Number of participants

data %>% 
  pull(id) %>% 
  unique %>% 
  length()

# Number of waves

data %>% 
  pull(year) %>% 
  unique()

data %>% 
  group_by(year) %>% 
  summarize(mean_srh = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_srh)) +
  geom_line() +
  geom_point() +
  labs(title = "Self-Rated Health Over 4 Waves",
       x = "Panel Wave Year",
       y = "Mean Self-Rated Health")

data %>% 
  group_by(year) %>% 
  ggplot(aes(x = health)) +
  geom_histogram() +
  facet_wrap(~ year) +
  labs(title = "Histogram of Self-Rated Health in 4 Waves")

data <- data %>% 
  filter(year != 2008)

```

There are `r data %>% pull(id) %>% unique %>% length()` participants who were surveyed in 5 waves of interest: years 2002, 2010, 2015, 2008, 2017. Self-rated health overall seems to have decreased over time, although not at a constant rate. 

# a) OLS

1. (a) Run an OLS regression, including at least one independent variable and a time variable (as dummies).  Explain how you think your independent variable relates to your dependent variable.  Interpret your results.  Did you find what you expected to find?  

```{r}

library(plm)

# OLS regression
summary(lm(health ~ watched_tv + as.factor(year), data))

eq.pool_tv <- plm(health ~ watched_tv + as.factor(year), # model formula
               index = c("id", "year"), # id & time variables
               model = "pooling", 
               data = data) ## this is equivalent to above OLS ##

summary(eq.pool_tv)

```

# b) Fixed Effects

(b) Then run a fixed effect model version of that OLS model.  Interpret your results.  Did you find what you expected to find?  Why?  Why not? 

```{r}

eq.fe_tv <- plm(health ~ watched_tv + as.factor(year), # model formula
             index = c("id", "year"), # id & time variables
             model = "within", 
             data = data) ## fixed effects ##

summary(eq.fe_tv)

```

# Additional predictor

(c) Then include an additional predictor in your fixed effects model that you think might account for the initial relationship you found between your X and your Y.  What effect does that new independent variable have in your new regression?

```{r}

eq.fe_tv_happy <- plm(health ~ watched_tv + felt_happy + as.factor(year), # model formula
             index = c("id", "year"), # id & time variables
             model = "within", 
             data = data) ## fixed effects ##

summary(eq.fe_tv_happy)

```

# d) Random Effects

(d) Then run a random effects model equivalent to your fixed effects model in step (b).  Interpret the results.

```{r}

eq.re_tv <- plm(health ~ watched_tv + as.factor(year), # model formula
             index = c("id", "year"), # id & time variables
             model = "random", 
             data = data) ## random effects ##

summary(eq.re_tv)

```

# e) Hausman test

(e) Run a Hausman test to compare your fixed effects and your random effects models.  What do you conclude? 

```{r}
library(stargazer)

phtest(eq.fe_tv, eq.re_tv) ## Hausman test comparing RE and FE ##

stargazer(eq.pool_tv, eq.fe_tv, eq.re_tv, 
          title="Regression Results", 
          align=TRUE, 
          dep.var.labels=c("Equalize"), 
       #   covariate.labels=c("Bills","2010"),  
          no.space=TRUE, 
          column.labels=c("Pooled", 
                        #  "First Diff", 
                          "Fixed Effects", "Random Effects"), 
          dep.var.caption="", 
          model.numbers=FALSE,
          type = "text", omit = "Constant")

```


# Additional Analyses

## OLS and FE for the additional variables

```{r}


# OLS - felt happy
eq.pool_happy <- plm(health ~ felt_happy + as.factor(year), # model formula
               index = c("id", "year"), # id & time variables
               model = "pooling", 
               data = data) 
summary(eq.pool_happy)

# FE - felt happy
eq.fe_happy <- plm(health ~ felt_happy + as.factor(year), # model formula
             index = c("id", "year"), # id & time variables
             model = "within", 
             data = data) ## fixed effects ##
summary(eq.fe_happy)

# OLS - depressed
eq.pool_depressed <- plm(health ~ depressed + as.factor(year), # model formula
               index = c("id", "year"), # id & time variables
               model = "pooling", 
               data = data) 

summary(eq.pool_depressed)

# FE - depressed
eq.fe_depressed <- plm(health ~ depressed + as.factor(year), # model formula
               index = c("id", "year"), # id & time variables
               model = "within", 
               data = data) 

summary(eq.fe_depressed)


```


## FE for different combinations of predictor variables

```{r}


eq.fe_happy_depressed <- plm(health ~ felt_happy + depressed + as.factor(year), # model formula
             index = c("id", "year"), # id & time variables
             model = "within", 
             data = data) ## fixed effects ##
summary(eq.fe_happy_depressed)

eq.fe_tv_happy_depressed <- plm(health ~ watched_tv + felt_happy + depressed + as.factor(year), # model formula
             index = c("id", "year"), # id & time variables
             model = "within", 
             data = data) ## fixed effects ##
summary(eq.fe_tv_happy_depressed)



```


## Comparison

```{r}

stargazer(eq.fe_tv_happy, eq.fe_tv_depressed, eq.fe_happy_depressed, #eq.fe_tv_happy_depressed,
          title="FE Results", 
          align=TRUE, 
          dep.var.labels=c("Equalize"), 
       #   covariate.labels=c("Bills","2010"),  
          no.space=TRUE, 
          column.labels=c("FE TV + Happy", 
                          "FE TV + Depressed", 
                          "FE Depressed + Happy" #,
                         # "FE TV + Depressed + Happy"
                          ), 
          dep.var.caption="", 
          model.numbers=FALSE,
          type = "text", omit = "Constant")

```


