---
title: "Self-Rated Health - World Values Survey"
author: "Christine Lucille Kuryla"
date: "2024-10-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
```

We have seen pattens in self-rated health in the GSS data. Let's explore the WVS data, EVS data, and IVS (WVS + IVS) to see if the pattern appears there as well. 

WVS data
* https://www.worldvaluessurvey.org/
* 7 survey waves (every 5 years)

EVS data
* https://search.gesis.org/research_data/ZA7503?doi=10.4232/1.14021 
* 5 survey waves (every 9 years) 

Integrated Values Survey (IVS)
* https://www.worldvaluessurvey.org/WVSEVStrend.jsp 
* EVS and WVS time-series data-sets are released independently by EVS/GESIS and the WVSA/JDS. The Integrated Values Surveys (IVS) dataset 1981-2022 can be constructed by merging the EVS Trend File 1981-2017 (doi:10.4232/1.14021) and the WVS trend 1981-2022 data-set doi:10.14281/18241.27. It is based on the Common EVS/WVS Dictionary (2021) and includes 452 surveys from 115 countries/territories.

# WVS Variables

<details>
  <summary>**Click to expand and explore the variables of interest, including the coding and the way the question is phrased.**</summary>

S012.- Date interview
Date of interview YYYYMMDD
(S012)
-3 Not applicable
-4 Not asked in survey

S020.- Year survey
Year of survey YYYY
(S020)
1981 1981

A008.- Feeling of happiness
Taking all things together, would you say you are:
(A008)
1 Very happy
2 Quite happy
3 Not very happy
4 Not at all happy
-1 Don't know
-2 No answer
-4 Not asked
-5 Missing; Not available

A009.- State of health (subjective)
All in all, how would you describe your state of health these days? Would you say it is...
(A009)
1 Very good
2 Good
3 Fair
4 Poor
5 Very poor
-1 Don't know
-2 No answer
-4 Not asked in survey
-5 Missing; Unknown

X003.- Age
Age
WVS Time Series 1981 2022
Variables Report V5.0
- 209 -
(X003)
-1 Don't know
-2 No answer
-3 Not applicable
-4 Not asked in survey
-5 Missing; Unknown

X001.- Sex
Sex
(X001)
1 Male
2 Female
-1 Don't know
-2 No answer
-4 Not asked
-5 Missing; Unknown

X002.- Year of birth
Can you tell me your year of birth, please?
(X002)
-1 Don't know
-2 No answer
-3 Not applicable
-4 Not asked in survey
-5 Missing; Unknown

X002_02A.- Respondents country of birth - ISO 3166-1 code
In which country were you born?
(X002_02A)
See annex number 21

S001.- Study
Study
(S001)
1 EVS
2 WVS
S002VS.- Chronology of EVS-WVS waves
Wave
(S002VS)
1 1981-1984
2 1989-1993
3 1994-1998
4 1999-2004
5 2005-2009
6 2010-2014
7 2017-2022
S003.- ISO 3166-1 numeric country code
Country code ISO 3166
(S003)
See annex number 1
COUNTRY_ALPHA.- ISO 3166-1 alpha-3 country code
(COUNTRY_ALPHA)
See annex number 2
COW_NUM.- CoW country code numeric
Country code CoW numeric
(COW_NUM)
See annex number 3
COW_ALPHA.- CoW country code alpha
CoW country code alpha
(COW_ALPHA)
See annex number 4

</details>

# Common variables

A009	State of health (subjective)
A008	Feeling of happiness
S020	Year survey
X002	Year of birth
X003	Age
X001	Sex
X051	Ethnic group
E033	Self positioning in political scale
B008	Protecting environment vs. Economic growth
S012	Date interview [YYYYMMDD]

S003	Country (ISO 3166-1 Numeric code)
COW_ALPHA	CoW country code alpha
COW_NUM	Country (CoW Numeric code)

A165	Most people can be trusted
A170	Satisfaction with your life
A173	How much freedom of choice and control

X002_02	Respondent born in [country]
X002_02A	Respondents country of birth: ISO 3166-1 code
X002_02B	Respondents country of birth: ISO 3166-1/3 Alpha code
X002_03	Year in which respondent came to live in [country]
X003R	Age recoded
X003R2	Age recoded (3 intervals)






# IVS

# Import, clean, format data from previous files

```{r redo_gss_analysis}
data_wvs_usa <- read_csv("data/wvs_usa.csv") %>% 
  filter(health %in% c(1, 2, 3, 4, 5))

data_evs_usa <- read_csv("data/evs_usa.csv") %>% 
  filter(health %in% c(1, 2, 3, 4, 5))

table(data_evs_usa$happy)
# CHECK THE CODING
table(data_wvs_usa$happy)
# CHECK THE CODING

hist(data_wvs_usa$health)
hist(data_wvs_usa$age)

hist(data_evs_usa$health)
hist(data_evs_usa$age)

evs_select <- data_evs_usa %>% 
                             select(age, year, cohort, sex, health, happy) 
wvs_select <- data_wvs_usa %>% 
                             select(age, year, cohort, sex, health, happy)

data_ivs_usa <- rbind(evs_select, wvs_select) %>% 
  filter(cohort > 1700) %>% 
  filter(sex %in% c(1,2)) %>% 
  filter(happy %in% c(1, 2, 3, 4)) %>%  # check the coding!
  filter(health %in% c(1, 2, 3, 4, 5))

write_csv(data_ivs_usa, "data/ivs_usa_selected.csv")
```


```{r}

data_ivs_usa <- read_csv("data/ivs_usa_selected.csv")

hist(data_ivs_usa$age)
hist(data_ivs_usa$year)
hist(data_ivs_usa$cohort)
hist(data_ivs_usa$sex)
hist(data_ivs_usa$health)
hist(data_ivs_usa$happy)


```



# IVS (WVS + EVS) Replicate GSS Analysis for New Data
```{r redo_gss_analysis_ivs}


data_ivs_usa %>% 
  filter(age > 18, age < 90) %>% 
  mutate(age = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = age)) +
  geom_line() +
  labs(title = "Average SRH Per Year for Each Age Group",
       subtitle = "IVS Dataset",
       y = "Average SRH", 
       x = "Year",
       color = "Age Group") +
  theme_minimal() +
  geom_point() 


# health vs age per year
data_ivs_usa %>% 
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)",
       subtitle = "IVS Dataset",
       y = "Average SRH", 
       x = "Age of Respondent")


# Aggregate slopes

# years_of_gss <- c(data_gss %>% select(year) %>% unique() )
# lm_health_v_age_0 <- data_gss %>%
#   group_by(year) %>%
#   summarize(coef = coef(lm(health ~ age, data = cur_data()))["age"])

# Perform linear regression for each year and extract the coefficient of 'age' with confidence intervals, se, t stat, p val
lm_health_v_age_0 <- data_ivs_usa %>%
  group_by(year) %>%
  do(tidy(lm(health ~ age, data = .), conf.int = TRUE)) %>%  # Add conf.int = TRUE for CIs
  filter(term == "age") %>%
  select(year, coef = estimate, conf.low, conf.high, se = std.error, t_statistic = statistic,  p_value = p.value)

# View the results with confidence intervals, se, t statistic, and p value
# print(lm_health_v_age_0)
knitr::kable(lm_health_v_age_0,
             caption = "IVS Dataset")

# Plot coefficients
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                 position=position_dodge(0.05)) +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    subtitle = "IVS Dataset",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Plot coefficients with CI
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                 position=position_dodge(0.05)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Add shaded area for confidence intervals
  labs(
    title = "Change in 'Age' Coefficient Over Years with Confidence Intervals",
    subtitle = "IVS Dataset",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Perform linear regression of 'coef' (age coefficient) vs 'year'
lm_coef_vs_year <- lm(coef ~ year, data = lm_health_v_age_0)

# View the summary of the regression
summary(lm_coef_vs_year)

ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                 position=position_dodge(0.05)) +
  geom_smooth(method = "lm", se = TRUE) +  # Adds the regression line with standard error shading
#  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Confidence intervals for the coefficients
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    subtitle = "IVS Dataset",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

data_ivs_usa %>% 
  filter(cohort > 1800, cohort < 2020) %>% 
  mutate(cohort = cut(cohort, breaks = 4)) %>% # Create cohorts with 6 breaks
  group_by(age, cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health, color = cohort)) +
  labs(title = "Age Profiles by Cohort", 
       subtitle = "IVS Dataset") +
  geom_line()



```


# Load data, wrangle, and recode data

## WVS

```{r load_data, eval = FALSE}

# Registration is needed to download the files, so we're just going to have to load them from my local computer

data_wvs_all <- read_csv("big_data/WVS_Time_Series_1981-2022_csv_v5_0.csv")

colnames()

data_wvs <- data_wvs_all %>% 
#  filter(S003 == 840) %>%   # United States 
  select(S003,          #.- ISO 3166-1 numeric country code
         COUNTRY_ALPHA, #.- ISO 3166-1 alpha-3 country code
         COW_NUM,       #.- CoW country code numeric
         COW_ALPHA,     # .- CoW country code alpha
         S012,          # date of interview
         S001,          # Study -- 1 EVS, 2 WVS
         S022,
         S023,
         X003, # age
         X001, # sex
         X002, # birth year
         S020, # year of survey
         A009, # self-rated health
         A008  # self-rated happiness
         ) %>% 
  mutate(year = as.numeric(str_sub(as.character(S022), start = 1, end = 4))) %>% 
  mutate(year2 = as.numeric(str_sub(as.character(S023), start = 1, end = 4))) 

data_wvs1 <- data_wvs %>% 
    mutate(country_code = S003) %>% 
    mutate(age = X003) %>% 
    mutate(sex = X001) %>% 
    mutate(cohort = X002) %>% 
    mutate(year = as.numeric(str_sub(as.character(S022), start = 1, end = 4))) %>%
    mutate(health = A009) %>% 
    mutate(happy = A008) %>% 
    mutate(year = as.numeric(str_sub(as.character(S022), start = 1, end = 4))) %>% 
    mutate(year2 = as.numeric(str_sub(as.character(S023), start = 1, end = 4))) %>% 
    mutate(health = 6 - health) %>% 
    mutate(happy = 5 - happy) #%>% 
   # select(age, cohort, year, health, happy, sex, country_code)

table(data_wvs1$year)

data_wvs_usa <- data_wvs1 %>% filter(COUNTRY_ALPHA == "USA")
  
  filter(S003 == 840) |   
         COUNTRY_ALPHA == "USA" |
         COW_NUM == 2 |
         COW_ALPHA == "USA"
           )

table(data_wvs_usa$year)

write_csv(data_wvs_usa, file = "data/wvs_usa.csv")

```


## EVS

```{r evs_data, eval = FALSE}

library(haven)

data_evs_all <- read_dta("big_data/EVS/ZA7503_v3-0-0.dta")

data_evs_selected <- data_evs_all %>% 
  select(A009,  	# State of health (subjective)
        A008,  	#	Feeling of happiness
        S020,  	#	Year survey
        X002,  	#	Year of birth
        X003,  	#	Age
        X001,  	#	Sex
        X051,  	#	Ethnic group
        E033,  	#	Self positioning in political scale
        B008,  	#	Protecting environment vs. Economic growth
        S012,  	#	Date interview [YYYYMMDD]
        S003,  	#	Country (ISO 3166-1 Numeric code)
   #     COW_ALPHA,  	#	CoW country code alpha
        COW_NUM,  	#	Country (CoW Numeric code)
        A165,  	#	Most people can be trusted
        A170,  	#	Satisfaction with your life
        A173,  	#	How much freedom of choice and control
        X002_02,  	#	Respondent born in [country]
        X002_02A,  	#	Respondents country of birth: ISO 3166-1 code
        X002_02B,  	#	Respondents country of birth: ISO 3166-1/3 Alpha code
        X002_03,  	#	Year in which respondent came to live in [country]
        X003R,  	#	Age recoded
        X003R2  	#	Age recoded (3 intervals)
  )

data_evs_recoded <- data_evs_selected %>% 
    mutate(country_code = S003) %>% 
    mutate(age = X003) %>% 
    mutate(sex = X001) %>% 
    mutate(cohort = X002) %>% 
#    mutate(year = as.numeric(str_sub(as.character(S022), start = 1, end = 4))) %>%
    mutate(health = A009) %>% 
    mutate(happy = A008) %>% 
  #  mutate(year = as.numeric(str_sub(as.character(S022), start = 1, end = 4))) %>% 
 #   mutate(year = as.numeric(str_sub(as.character(S012), start = 1, end = 4))) %>% 
  #  mutate(year2 = as.numeric(str_sub(as.character(S023), start = 1, end = 4))) %>% 
    mutate(health = 6 - health) %>% 
    mutate(happy = 5 - happy) %>% 
    mutate(year = S020)  	#	Year survey #%>% 
   # select(age, cohort, year, health, happy, sex, country_code)

table(data_evs_recoded$year)

data_evs_usa <- data_evs_recoded %>% 
  filter(country_code == 840 |   
       #  COUNTRY_ALPHA == "USA" |
         COW_NUM == 2
       #  COW_ALPHA == "USA"
         )

data_evs_world <- data_evs_recoded

table(data_evs_usa$year)

table(data_evs_world$year)

write_csv(data_evs_world, file = "data/data_evs_world.csv")


write_csv(data_evs_usa, "data/evs_usa.csv")

```


# Redo GSS analysis for WVS

```{r redo_gss_analysis_wvs, eval = FALSE}
data_wvs_usa %>% 
  filter(age > 18, age < 90) %>% 
  mutate(age = cut(age, breaks = 8)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = age)) +
  geom_line() +
  geom_point()


# health vs age per year
data_wvs_usa %>% 
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)" )



# Aggregate slopes

# years_of_gss <- c(data_gss %>% select(year) %>% unique() )
# lm_health_v_age_0 <- data_gss %>%
#   group_by(year) %>%
#   summarize(coef = coef(lm(health ~ age, data = cur_data()))["age"])

# Perform linear regression for each year and extract the coefficient of 'age' with confidence intervals, se, t stat, p val
lm_health_v_age_0 <- data_wvs_usa %>%
  group_by(year) %>%
  do(tidy(lm(health ~ age, data = .), conf.int = TRUE)) %>%  # Add conf.int = TRUE for CIs
  filter(term == "age") %>%
  select(year, coef = estimate, conf.low, conf.high, se = std.error, t_statistic = statistic,  p_value = p.value)

# View the results with confidence intervals, se, t statistic, and p value
# print(lm_health_v_age_0)
knitr::kable(lm_health_v_age_0)

# Plot coefficients
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                 position=position_dodge(0.05)) +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Plot coefficients with CI
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_line() +
  geom_point() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Add shaded area for confidence intervals
  labs(
    title = "Change in 'Age' Coefficient Over Years with Confidence Intervals",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Perform linear regression of 'coef' (age coefficient) vs 'year'
lm_coef_vs_year <- lm(coef ~ year, data = lm_health_v_age_0)

# View the summary of the regression
summary(lm_coef_vs_year)

ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                 position=position_dodge(0.05)) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +  # Adds the regression line with standard error shading
#  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Confidence intervals for the coefficients
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

data_wvs_usa %>% 
  filter(cohort > 1800, cohort < 2020) %>% 
  mutate(cohort = cut(cohort, breaks = 4)) %>% # Create cohorts with 6 breaks
  group_by(age, cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health, color = cohort)) +
  labs(title = "Age Profiles by Cohort") +
  geom_line()





```

