---
title: "Self-Rated Health - World Values Survey"
author: "Christine Lucille Kuryla"
date: "2024-10-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

We have seen pattens in self-rated health in the GSS data. Let's explore the WVS data (https://www.worldvaluessurvey.org/) to see if the pattern appears there as well. 

# Variables

S012.- Date interview
Date of interview YYYYMMDD
(S012)
-3 Not applicable
-4 Not asked in survey

S020.- Year survey
Year of survey YYYY
(S020)
1981 1981

A008.- Feeling of happiness
Taking all things together, would you say you are:
(A008)
1 Very happy
2 Quite happy
3 Not very happy
4 Not at all happy
-1 Don't know
-2 No answer
-4 Not asked
-5 Missing; Not available

A009.- State of health (subjective)
All in all, how would you describe your state of health these days? Would you say it is...
(A009)
1 Very good
2 Good
3 Fair
4 Poor
5 Very poor
-1 Don't know
-2 No answer
-4 Not asked in survey
-5 Missing; Unknown

X003.- Age
Age
WVS Time Series 1981 2022
Variables Report V5.0
- 209 -
(X003)
-1 Don't know
-2 No answer
-3 Not applicable
-4 Not asked in survey
-5 Missing; Unknown

X001.- Sex
Sex
(X001)
1 Male
2 Female
-1 Don't know
-2 No answer
-4 Not asked
-5 Missing; Unknown

X002.- Year of birth
Can you tell me your year of birth, please?
(X002)
-1 Don't know
-2 No answer
-3 Not applicable
-4 Not asked in survey
-5 Missing; Unknown

X002_02A.- Respondents country of birth - ISO 3166-1 code
In which country were you born?
(X002_02A)
See annex number 21

S001.- Study
Study
(S001)
1 EVS
2 WVS
S002VS.- Chronology of EVS-WVS waves
Wave
(S002VS)
1 1981-1984
2 1989-1993
3 1994-1998
4 1999-2004
5 2005-2009
6 2010-2014
7 2017-2022
S003.- ISO 3166-1 numeric country code
Country code ISO 3166
(S003)
See annex number 1
COUNTRY_ALPHA.- ISO 3166-1 alpha-3 country code
(COUNTRY_ALPHA)
See annex number 2
COW_NUM.- CoW country code numeric
Country code CoW numeric
(COW_NUM)
See annex number 3
COW_ALPHA.- CoW country code alpha
CoW country code alpha
(COW_ALPHA)
See annex number 4

```{r load_data, eval = FALSE}

# Registration is needed to download the files, so we're just going to have to load them from my local computer

data_wvs_all <- read_csv("big_data/WVS_Time_Series_1981-2022_csv_v5_0.csv")

colnames()

data_wvs <- data_wvs_all %>% 
#  filter(S003 == 840) %>%   # United States 
  select(S003,          #.- ISO 3166-1 numeric country code
         COUNTRY_ALPHA, #.- ISO 3166-1 alpha-3 country code
         COW_NUM,       #.- CoW country code numeric
         COW_ALPHA,     # .- CoW country code alpha
         S012,          # date of interview
         S001,          # Study -- 1 EVS, 2 WVS
         S022,
         S023,
         X003, # age
         X001, # sex
         X002, # birth year
         S020, # year of survey
         A009, # self-rated health
         A008  # self-rated happiness
         ) %>% 
  mutate(year = as.numeric(str_sub(as.character(S022), start = 1, end = 4))) %>% 
  mutate(year2 = as.numeric(str_sub(as.character(S023), start = 1, end = 4))) 

data_wvs1 <- data_wvs %>% 
    mutate(country_code = S003) %>% 
    mutate(age = X003) %>% 
    mutate(sex = X001) %>% 
    mutate(cohort = X002) %>% 
    mutate(year = as.numeric(str_sub(as.character(S022), start = 1, end = 4))) %>% 
    mutate(health = A009) %>% 
    mutate(happy = A008) %>% 
    mutate(health = 6 - health) %>% 
    mutate(happy = 5 - happy) #%>% 
   # select(age, cohort, year, health, happy, sex, country_code)

table(data_wvs1$year)

data_wvs_usa <- data_wvs1 %>% filter(COUNTRY_ALPHA == "USA")
  
  filter(S003 == 840) |   
         COUNTRY_ALPHA == "USA" |
         COW_NUM == 2 |
         COW_ALPHA == "USA"
           )

table(data_wvs_usa$year)

write_csv(data_wvs_usa, file = "data/wvs_usa.csv")

```

```{r redo_gss_analysis}

data_wvs_usa <- read_csv("data/wvs_usa.csv") %>% 
  filter(health %in% c(1, 2, 3, 4, 5))

hist(data_wvs_usa$health)
hist(data_wvs_usa$age)

data_wvs_usa %>% 
  filter(age > 18, age < 90) %>% 
  mutate(age = cut(age, breaks = 8)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = age)) +
  geom_line() +
  geom_point()


# health vs age per year
data_wvs_usa %>% 
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)" )

library(broom)

# Aggregate slopes

# years_of_gss <- c(data_gss %>% select(year) %>% unique() )
# lm_health_v_age_0 <- data_gss %>%
#   group_by(year) %>%
#   summarize(coef = coef(lm(health ~ age, data = cur_data()))["age"])

# Perform linear regression for each year and extract the coefficient of 'age' with confidence intervals, se, t stat, p val
lm_health_v_age_0 <- data_wvs_usa %>%
  group_by(year) %>%
  do(tidy(lm(health ~ age, data = .), conf.int = TRUE)) %>%  # Add conf.int = TRUE for CIs
  filter(term == "age") %>%
  select(year, coef = estimate, conf.low, conf.high, se = std.error, t_statistic = statistic,  p_value = p.value)

# View the results with confidence intervals, se, t statistic, and p value
# print(lm_health_v_age_0)
knitr::kable(lm_health_v_age_0)

# Plot coefficients
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Plot coefficients with CI
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_line() +
  geom_point() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Add shaded area for confidence intervals
  labs(
    title = "Change in 'Age' Coefficient Over Years with Confidence Intervals",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Perform linear regression of 'coef' (age coefficient) vs 'year'
lm_coef_vs_year <- lm(coef ~ year, data = lm_health_v_age_0)

# View the summary of the regression
summary(lm_coef_vs_year)

ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  # Adds the regression line with standard error shading
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Confidence intervals for the coefficients
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

data_wvs_usa %>% 
  filter(cohort > 1800, cohort < 2020) %>% 
  mutate(cohort = cut(cohort, breaks = 4)) %>% # Create cohorts with 6 breaks
  group_by(age, cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health, color = cohort)) +
  labs(title = "Age Profiles by Cohort") +
  geom_line()

```


