---
title: 'Mortality and SRH: GSS 2014'
author: "Christine Lucille Kuryla"
date: "2025-01-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Context

Self-rated health (SRH) is often used because of its well-documented association with mortality, even beyond comorbidities and other covariates. Here, we explore whether the predictive power of SRH on mortality has changed over the years in the GSS dataset. 

We will use the General Social Survey (GSS). For the years 1978 - 2010, there is a variable `death` which is the possible vital status as of 2014. https://gssdataexplorer.norc.org/variables/7426/vshow 

https://chatgpt.com/share/6777c9b5-551c-8002-9603-e2d1856253e7

# Limitations

The optimal analysis here would be survival analysis, however, we do know know the year of death for each person, only whether they were alive or deceased in 2014. We must recognize the limitation of a single binary outcome (“Alive/Dead by 2014”). The simplest binary logistic approach (alive vs. dead by 2014) effectively treats someone interviewed in 1978 and followed for 36 years the same as someone interviewed in 2010 and followed for only 4 years. If the older wave has higher mortality simply because they’ve had more time to die, that might artificially inflate differences across survey years. Consequence: Period effect might get conflated with time at risk. We may see stronger SRH–mortality associations in older data simply because more events (deaths) could have occurred.


# Load data

```{r}

#data_gss <- read_csv(here("data/cleaned/gss_groups.csv")) %>% 

data_gss <- read_csv("data/extracted_gss_variables.csv") %>% 
  filter(cohort != 9999) %>% 
  filter(!(is.na(death))) %>% 
#  na.omit() %>% 
  mutate(health = 5 - health)  %>%  # reverse the coding so it's more intuitive (higher number for excellent, lower number for poor)
  mutate(happy = 4 - happy) %>% # same
  mutate(life = 4 - life) %>% # reverse again, these variables tend to be unintuitively ordered!!!
  mutate(satfin = 4 - satfin) %>% # same again!
    mutate(
    health_cat = factor(health, 
                        levels = 1:4,
                        labels = c("Poor", "Fair", "Good", "Excellent")),
    # period_cut_6 = as.factor(cut(data_gss$year, 6)),
    # period_cut_10 = as.factor(cut(data_gss$year, 10)),
    # period_cut_12 = as.factor(cut(data_gss$year, 12)),
    # period_groups = as.factor(cut(data_gss$year, 12)),
    period_7yr = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1982, 1990, 1998, 2006, 2014, Inf),
                            labels = c("1974-1982", "1983-1990", "1991-1998", 
                                       "1999-2006", "2007-2014", "2015-2022"),
                            right = TRUE
                            )
                          ),
    # period_5yr = as.factor(
    #                         cut(
    #                         year,
    #                         breaks = c(1973, 1978, 1984, 1989, 1994, 1999, 2004, 2009 ,2014, Inf),
    #                         labels = c("1974-1979", "1982-1990", "1990-1998", 
    #                                    "1998-2006", "2006-2014", "2014-2022"),
    #                         right = TRUE
    #                         )
    #                       ),
    period_decade = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1979, 1989, 1999, 2009, 2019, Inf),
                            labels = c("1974-1979", "1980-1989", "1990-1999",
                                       "2000-2009", "2010-2019", "2020-2024"),
                            right = TRUE
                            )
                          ),
    age_group = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_groups = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_group_small = as.factor( 
                            cut(
                              age,
                              breaks = c(seq(15, 75, by = 5), Inf),  # Define breaks up to 75 and include Inf for the last group
                              labels = c("16-20", "21-25", "26-30", "31-35", "36-40", "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76+"),
                              right = FALSE  # Makes intervals left-closed, i.e., [x, y)
                            )
                          )
                          ,
    generation_5total = factor(
      case_when(
        cohort >= 1901 & cohort <= 1927 ~ "Greatest (1901-1927)",
        cohort >= 1928 & cohort <= 1945 ~ "Silent (1928-1945)",
        cohort >= 1946 & cohort <= 1964 ~ "Boomers (1946-1964)",
        cohort >= 1965 & cohort <= 1980 ~ "Gen X (1965-1980)",
        # cohort >= 1981 & cohort <= 1996 ~ "Millennials (1981-1996)",
        # cohort >= 1997 & cohort <= 2012 ~ "Gen Z (1997-2012)",
        cohort >= 1981  ~ "Millennials / Gen Z (1981-2004)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest (1901-1927)",
        "Silent (1928-1945)",
        "Boomers (1946-1964)",
        "Gen X (1965-1980)",
        "Millennials / Gen Z (1981-2004)"
       # "Millennials (1981-1996)",
      #  "Gen Z (1997-2012)"#,
     #   "Other"
      )
    ),
    generation_10total = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1914 ~ "Greatest Early (1901-1914)",
        generation == "Greatest (1901-1927)" & cohort > 1914 ~ "Greatest Late (1915-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1936 ~ "Silent Early (1928-1936)",
        generation == "Silent (1928-1945)" & cohort > 1936 ~ "Silent Late (1937-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1955 ~ "Boomers Early (1946-1955)",
        generation == "Boomers (1946-1964)" & cohort > 1955 ~ "Boomers Late (1956-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1972 ~ "Gen X Early (1965-1972)",
        generation == "Gen X (1965-1980)" & cohort > 1972 ~ "Gen X Late (1973-1980)",
        cohort >= 1981 & cohort <= 1988 ~ "Millennials Early (1981-1988)",
        cohort > 1988 ~ "Millennials Late / Gen Z Early (1989-2004)",
        # generation == "Millennials (1981-1996)" & cohort > 1988 ~ "Millennials Late (1989-1996)",
        # generation == "Gen Z (1997-2012)" & cohort <= 2004 ~ "Gen Z Early (1997-2004)",
        # generation == "Gen Z (1997-2012)" & cohort > 2004 ~ "Gen Z Late (2005-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1914)", "Greatest Late (1915-1927)",
        "Silent Early (1928-1936)", "Silent Late (1937-1945)",
        "Boomers Early (1946-1955)", "Boomers Late (1956-1964)",
        "Gen X Early (1965-1972)", "Gen X Late (1973-1980)",
        "Millennials Early (1981-1988)", "Millennials Late / Gen Z Early (1989-2004)"
      #  "Millennials Early (1981-1988)", "Millennials Late (1989-1996)",
      #  "Gen Z Early (1997-2004)", "Gen Z Late (2005-2012)"#,
     #   "Other"
      )
    ),
    generation_15total = factor(
      case_when(
        cohort >= 1900 & cohort <= 1910 ~ "Greatest Early (1901-1910)",
        cohort >= 1911 & cohort <= 1918 ~ "Greatest Mid (1911-1918)",
        cohort >= 1919 & cohort <= 1927 ~ "Greatest Late (1919-1927)",
        cohort >= 1928 & cohort <= 1934 ~ "Silent Early (1928-1934)",
        cohort >= 1935 & cohort <= 1940 ~ "Silent Mid (1935-1940)",
        cohort >= 1941 & cohort <= 1945 ~ "Silent Late (1941-1945)",
        cohort >= 1945 & cohort <= 1951 ~ "Boomers Early (1946-1951)",
        cohort >= 1952 & cohort <= 1958 ~ "Boomers Mid (1952-1958)",
        cohort >= 1959 & cohort <= 1964 ~ "Boomers Late (1959-1964)",
        cohort >= 1965 & cohort <= 1970 ~ "Gen X Early (1965-1970)",
        cohort >= 1971 & cohort <= 1976 ~ "Gen X Mid (1971-1976)",
        cohort >= 1977 & cohort <= 1980 ~ "Gen X Late (1977-1980)",
        cohort >= 1981 & cohort <= 1986 ~ "Millennials Early (1981-1986)",
        cohort >= 1987 & cohort <= 1992 ~ "Millennials Mid (1987-1992)",
        cohort >= 1993 ~ "Millennials Late / Gen Z (1993-2004)",
    #    generation == "Gen Z (1997-2012)" & cohort <= 2002 ~ "Gen Z Early (1997-2002)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2002 & cohort <= 2008 ~ "Gen Z Mid (2003-2008)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2008 ~ "Gen Z Late (2009-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1910)", "Greatest Mid (1911-1918)", "Greatest Late (1919-1927)",
        "Silent Early (1928-1934)", "Silent Mid (1935-1940)", "Silent Late (1941-1945)",
        "Boomers Early (1946-1951)", "Boomers Mid (1952-1958)", "Boomers Late (1959-1964)",
        "Gen X Early (1965-1970)", "Gen X Mid (1971-1976)", "Gen X Late (1977-1980)",
        "Millennials Early (1981-1986)", "Millennials Mid (1987-1992)", 
        "Millennials Late / Gen Z (1993-2004)"
        #"Millennials Late (1993-1996)",
      #  "Gen Z Early (1997-2002)", "Gen Z Mid (2003-2008)", "Gen Z Late (2009-2012)" #,
      #  "Other"
      )
    )
  ) %>% 
  mutate(
    # 2-year periods
    period_2yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1975, 1977, 1979, 1981, 1983, 1985, 1987, 1989, 1991,
          1993, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011,
          2013, 2015, 2017, 2019, 2021, Inf
        ),
        labels = c(
          "1974-1975", "1976-1977", "1978-1979", "1980-1981", "1982-1983",
          "1984-1985", "1986-1987", "1988-1989", "1990-1991", "1992-1993",
          "1994-1995", "1996-1997", "1998-1999", "2000-2001", "2002-2003",
          "2004-2005", "2006-2007", "2008-2009", "2010-2011", "2012-2013",
          "2014-2015", "2016-2017", "2018-2019", "2020-2021", "2022-2022"
        ),
        right = TRUE
      )
    ),
    
    # 3-year periods
    period_3yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1976, 1979, 1982, 1985, 1988, 1991, 1994, 1997, 2000,
          2003, 2006, 2009, 2012, 2015, 2018, 2021, Inf
        ),
        labels = c(
          "1974-1976", "1977-1979", "1980-1982", "1983-1985",
          "1986-1988", "1989-1991", "1992-1994", "1995-1997",
          "1998-2000", "2001-2003", "2004-2006", "2007-2009",
          "2010-2012", "2013-2015", "2016-2018", "2019-2021", "2022-2022"
        ),
        right = TRUE
      )
    ),
    
    # 5-year periods
    period_5yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1978, 1983, 1988, 1993, 1998, 2003, 2008, 2013, 2018, Inf
        ),
        labels = c(
          "1974-1978", "1979-1983", "1984-1988", "1989-1993",
          "1994-1998", "1999-2003", "2004-2008", "2009-2013",
          "2014-2018", "2019-2022"
        ),
        right = TRUE
      )
    ),
    
    # 7-year periods
    period_7yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1980, 1987, 1994, 2001, 2008, 2015, Inf
        ),
        labels = c(
          "1974-1980", "1981-1987", "1988-1994",
          "1995-2001", "2002-2008", "2009-2015", "2016-2022"
        ),
        right = TRUE
      )
    ),
    
    # 8-year periods
    period_8yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1981, 1989, 1997, 2005, 2013, 2021, Inf
        ),
        labels = c(
          "1974-1981", "1982-1989", "1990-1997",
          "1998-2005", "2006-2013", "2014-2021", "2022-2022"
        ),
        right = TRUE
      )
    ),
    
    # 10-year periods
    period_10yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1983, 1993, 2003, 2013, Inf
        ),
        labels = c(
          "1974-1983", "1984-1993", "1994-2003",
          "2004-2013", "2014-2022"
        ),
        right = TRUE
      )
    )
  ) %>% 
  mutate(period_7total = period_7yr,
         period_10total = period_5yr,
         period_17total = period_3yr) %>% 
  mutate(
    # -----------------------
    # 6 AGE CATEGORIES
    # -----------------------
    age_6cat = as.factor(
      cut(
        age,
        # 7 breakpoints --> 6 intervals
        breaks = c(17, 30, 40, 50, 60, 70, 89),
        labels = c(
          "18-29",  # (17, 30]
          "30-39",  # (30, 40]
          "40-49",  # (40, 50]
          "50-59",  # (50, 60]
          "60-69",  # (60, 70]
          "70-89"   # (70, 89]
        ),
        right = TRUE
      )
    ),
    
    # -----------------------
    # 10 AGE CATEGORIES
    # -----------------------
    age_10cat = as.factor(
      cut(
        age,
        # 11 breakpoints --> 10 intervals
        breaks = c(17, 24, 31, 38, 45, 52, 59, 66, 73, 80, 89),
        labels = c(
          "18-24",  # (17, 24]
          "25-31",  # (24, 31]
          "32-38",  # (31, 38]
          "39-45",  # (38, 45]
          "46-52",  # (45, 52]
          "53-59",  # (52, 59]
          "60-66",  # (59, 66]
          "67-73",  # (66, 73]
          "74-80",  # (73, 80]
          "81-89"   # (80, 89]
        ),
        right = TRUE
      )
    ),
    
    # -----------------------
    # 16 AGE CATEGORIES
    # -----------------------
    age_16cat = as.factor(
      cut(
        age,
        # 17 breakpoints --> 16 intervals
        breaks = c(17, 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 85, 87, 89),
        labels = c(
          "18-22",  # (17, 22]
          "23-27",  # (22, 27]
          "28-32",  # (27, 32]
          "33-37",  # (32, 37]
          "38-42",  # (37, 42]
          "43-47",  # (42, 47]
          "48-52",  # (47, 52]
          "53-57",  # (52, 57]
          "58-62",  # (57, 62]
          "63-67",  # (62, 67]
          "68-72",  # (67, 72]
          "73-77",  # (72, 77]
          "78-82",  # (77, 82]
          "83-85",  # (82, 85]
          "86-87",  # (85, 87]
          "88-89"   # (87, 89]
        ),
        right = TRUE
      )
    )
    
  ) %>% 
    mutate(srh_num = health,
         srh_cat = health_cat) 

glimpse(data_gss)
summary(data_gss)

table(data_gss$generation_5total)
table(data_gss$generation_10total)
table(data_gss$generation_15total)

table(data_gss$period_7total)
table(data_gss$period_10total)
table(data_gss$period_17total)

table(data_gss$srh_num)

summary(data_gss$age)

```

### Add more variables

```{r}

data_gss <- data_gss %>% 
  mutate(id = row_number()) %>% 
  mutate(SRH = health) %>% 
  mutate(birth_year = cohort) %>% 
  died_by_2014 = 

```


# Method 1: Control for "years of exposure" or "time at risk"

